---
title: "Pre-Lab 1: Summarizing and Visualizing Data"
output: learnr::tutorial
tutorial:
  id: "Pre-Lab 1"
  version: 1.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
#install.packages(c('dplyr', 'ggplot2', 'learnr'))
#install.packages(c('learnr', 'shiny'))

knitr::opts_chunk$set(echo = FALSE)
titanic <- read.csv("http://calpoly.edu/~kbodwin/data/titanic0.csv")
titanic = titanic[,-1]
titanic_orig <- titanic
source("http://calpoly.edu/~kbodwin/scripts/packageHelpers.R")
source("http://calpoly.edu/~kbodwin/scripts/makeStrings.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

### %ni = new input (red), %oi = old input (blue)
```


## Reading Datasets

Today we'll be looking at a dataset containing information about passengers on the H.M.S. Titanic.


### Loading the data

The first step of any data analysis in **R** will be to "load" the data.  This means taking a dataset from online or from a folder on your computer and putting that information into the **R** program.

This dataset is already stored online, at [my personal website](http://calpoly.edu/~kbodwin/data/titanic0.csv) The information is stored in "csv" format, or "comma-separated values". We will simply use an **R** function called `read.csv()` to get this information and store it.

When you load a dataset, you will need to give it a name.  This is so that in the future, you can tell **R** where to find the information it needs - especially once you are dealing with more than one dataset.

Throughout this class, we will do our best to follow the [Tidyverse Style Guide](http://style.tidyverse.org/), which gives some guidelines for making your code look nice and easy to read.  

The style guide suggests giving names that use only lowercase letters and underscores ("_").  The name should also be descriptive; for example, for a dataset containing incomes of workers in the United States, we might go with `us_incomes_2018` rather than `incomes` or `my_dataset`.

We will name this dataset `titanic`.  The code to load the data looks like this:

```{r dataset_name}
# textInput("dat_name",
#           "What name would you like to give this dataset?",
#           value = "titanic",
#           width = '80%', placeholder = NULL)

```

```{r, context="server"}

output$code_1_1 <- renderText(
        makePrintText(
          base_string = "%ni <- read.csv('%oi')",
          old_input = c("http://calpoly.edu/~kbodwin/data/titanic0.csv"),
          new_input = c("titanic")
        )
  )

```

```{r}
htmlOutput("code_1_1")
```

You should read the `<-` symbol as "save the information on the right under the name on the left".  In this case, we are saving all of the data in the file `titanic0.csv` and giving it the name of our choosing, `titanic`.

### Checking out the data

For this lab, the dataset is already loaded for you.  The first thing we generally want to do is familiarize ourselves with the data. You can check out the first few rows of your dataset using the function `head()`, or you can look at some summary information about all the variables in the dataset using `summary()`.

When we want to do something to our dataset, we use a *pipe*, which looks like `%>%`.  You can think of this as "sending" the dataset into a process.

```{r, context="server"}

bob <- makePrintText(
          base_string = "%oi %>% head() \n %oi %>% summary()",
          old_input = c("titanic", "titanic")
        )

output$code_1_2 <- renderText(bob)

output$code_1_3 <- renderText(
        makePrintText(
          base_string = "head(%oi) \n summary(%oi)",
          old_input = c("titanic", "titanic")
        )
  )
```

```{r}
htmlOutput("code_1_2")
#codestring <- paste(codestring, code_1_2, sep = "\n")
```

Time to try it!  Edit the code in the box below, then click "Run Code".

```{r ex1.2, exercise = TRUE}
head()
summary()
```

Note that we can also simply put the name `titanic` inside the parentheses of the function, like this:

```{r}
htmlOutput("code_1_3")
```

However, it is good to get into the habit of using the pipe - as we will see a bit later on, this will become convenient when we want to use many functions in a row on the same dataset.

### Questions:

```{r q_1.1}
textInput("a_1.1.1",
           "How many Titanic passengers are represented in this dataset?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.1")

textInput("a_1.1.2",
           "What was the median age of passengers in this dataset?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.2")

textInput("a_1.1.3",
           "What was the mean cost of a ticket on the Titanic, based on this dataset?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.3")
```

```{r, context = "server"}

output$check_1.1.1 <- renderText(
  if(input$a_1.1.1 == dim(titanic)[1]){
      correct
  }else if(input$a_1.1.1 == ""){
     ""
  }else{
    incorrect
  }
)

output$check_1.1.2 <- renderText(
  if(input$a_1.1.2 == median(titanic$Age, na.rm = TRUE)){
      correct
  }else if(input$a_1.1.2 == ""){
     ""
  }else{
    incorrect
  }
)

output$check_1.1.3 <- renderText(
  if(input$a_1.1.3 == mean(titanic$Fare, na.rm = TRUE)){
      correct
  }else if(input$a_1.1.3 == ""){
     ""
  }else{
    incorrect
  }
)


```

## Cleaning Datasets

It is very rare that data comes to us in exactly the format we hope, or that **R** interprets the variable exactly as it should.  You will nearly always need to clean the data a bit before you begin calculating or plotting anything.  We will practice with one of the common issues today.

### Missing Data

Sometimes, values in the dataset are not present; perhaps because they were misrecorded or unavailable.

```{r q_2_1}
textInput("a_2_1",
           "In our current titanic dataset, which variables have missing values?",
           value = "",
           width = '100%', 
           placeholder = NULL)

htmlOutput("check_2.1")
```

```{r, context = "server"}

output$check_2.1 <- renderText(
  if(grepl("Fare", input$a_2_1, ignore.case = TRUE) & grepl("Age", input$a_2_1, ignore.case = TRUE)){
      correct
  }else if(input$a_2_1 == ""){
     ""
  }else{
    incorrect
  }
)


```

You, as the scientist, will need to decide how to handle these missing values.  This will depend on your **research question** and your choice of **analysis**.  For now, we will simply toss out the rows (passengers) for which we are missing some information.

The function `na.omit()` will automatically drop all rows from a dataset that have any values of `NA` in them.

In the code box below, pipe `titanic` into `na.omit()` and run it.

```{r ex_2_1, exercise = TRUE}

```

Suppose we want to take a look at our dataset once again, without those pesky `NA`'s.

```{r, echo = TRUE}
titanic %>% summary()
```

Oh no!  They are still there!  What happened?

We have fallen into a common mistake when using the `%>%` approach: we did not actually save our updated dataset.  That is, we told **R**, "Please do `na.omit()` to `titanic`", but we did not use `<-` to tell it "... and save the results under this name".

Although we *could* come up with another name - perhaps `titanic2` - and create a second dataset to work with, that isn't really necessary.  Most of the time, we will simply save our updated data with the same name as the old one.

While convenient, this updating can be dangerous.  We have now **replaced** the dataset `titanic` with a new version.  If we made a mistake, our original data is gone!

If this happens at home, you can always simply reload your data by running `read.csv()` again. For the purposes of this lab, a copy of the original dataset called `titanic_orig` is already loaded into **R**.  Any time you want to "reset" your data, simply run the line

```{r, eval = FALSE, echo = TRUE}
titanic <- titanic_orig
```

in one of your code boxes.

### Question:

Finish the code below, using piping and `na.omit()`, to update the dataset.

```{r ex_2_2, exercise = TRUE}
titanic <- 
```

```{r}

# htmlOutput("check_2.2")
```

```{r, context = "server"}
# 
# output$check_2.2 <- renderText(
#   if(sum(is.na(titanic)) == 0){
#       congrats
#   }else{
#      ""
#   }
# )


```


## Adjusting Variables

Next we may want to alter some of our variables in our dataset, to better represent what we wish to study.

### Quantitative and Categorical variables

Notice that when you ran `summary(titanic)`, you were given the mean and 5-number summary of the variable called `Pclass`.  This is because the information for that variable is given as number corresponding to the cabin class of the passenger (1, 2, or 3).  When **R** encounters numbers, it treats the variable as *quantitative*.  We know, however, that this should be *categorical* and *ordinal*.

To fix this, we will `mutate()` our dataset, which means adding or changing a variable.  The function `factor()` will turn a numeric variable into categories.


```{r factor_var}
textInput("pclass",
          "What is the name of the existing variable that you would like to make categorical?",
          value = "",
          width = '80%', placeholder = NULL)


textInput("new_pclass",
          "What name would you like to give the new version of the variable?",
          value = "",
          width = '80%', placeholder = NULL)


```

```{r, context="server"}

output$code_3_1 <- renderText(
        makePrintText(
          base_string = "%oi <- %oi %>% mutate( \n \t %ni = factor(%ni) \n )",
          old_input = c("titanic", "titanic"),
          new_input = c(input$new_pclass, input$pclass)
        )
  )

```

The code for this will look like:

```{r}
htmlOutput("code_3_1")
```

Run the code yourself, and check that the output shows a variable with your chosen name.

```{r ex_3_1, exercise = TRUE}
titanic %>% mutate(
  
)
```

Perhaps you think it might be convenient to have the values of your new variable be more informative than `0` and `1`.  This is easy to do with `factor()`.

```{r factor_var_names}
textInput("word_1",
          "What label would you like to use for 1?",
          value = "",
          width = '80%', placeholder = NULL)


textInput("word_2",
          "What label would you like to use for 2?",
          value = "",
          width = '80%', placeholder = NULL)

textInput("word_3",
          "What label would you like to use for 3?",
          value = "",
          width = '80%', placeholder = NULL)


```

```{r, context="server"}

output$code_3_2 <- renderText(  
  makePrintText(
          base_string = "titanic <- titanic %>% mutate( \n \t %oi = factor(%oi, \n \t \t levels = c(1,2,3), \n \t \t labels = c('%ni', '%ni', '%ni')) \n )",
          old_input = c(input$new_pclass, input$pclass),
          new_input = c(input$word_1, input$word_2, input$word_3)
        ))
```

```{r}
htmlOutput("code_3_2")
```

Run the code, and check that it did what you hoped it would.

```{r ex_3_2, exercise = TRUE}

```

This is your first experience with optional arguments in a function!  Notice that `factor()` worked just fine with only a variable name, but that you were able to do even more with the options `levels` and `labels`.


## Visualizing Categorical Variables

Data loaded?  Check!  Data cleaned?  Check!  Time for the fun stuff.  Let's begin by making some nice plots for the *categorical* variables in our dataset.


### Bar Plots

We will be using a package called `ggplot2` to make all of our visualizations.  This package has a very specific format for plotting.  

First, we tell **R** that we want to make a plot from a certain dataset, and we specify the **aethetic**, `aes()`, which determines the variables we want to visualize.  

Then, we add the type of plot we want, and any extra options.

```{r, echo=FALSE}
textInput("plot_var_1", 
          "What variable would you like to plot?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_dat_1", 
          "What dataset is that variable found in?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

bs_4_1 = reactive("ggplot(%ni, aes(x = %ni)) + geom_bar()")

htmlOutput("code_4_1")

plotOutput("plot_4_1")
```


```{r, context="server"}
output$plot_4_1 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_4_1(), 
      new_input = c(input$plot_dat_1, input$plot_var_1)
    )
   ))
})

output$code_4_1 <- renderText(
        makePrintText( 
          base_string = bs_4_1(), 
          new_input = c(input$plot_dat_1, input$plot_var_1)
        )
  )

```

Beyond the crucial information of the dataset and the variable(s) to plot, there are also plenty of optional settings to make our image nicer.

```{r, echo=FALSE}
textInput("plot_title_1", 
          "Title of the plot:", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_xlab_1", 
          "Label for x-axis:", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

radioButtons("fill_1", 
             "Should the bars be different colors?",
                   choices = c("No" = "",
                               "Yes" = ", fill = %oi")
              )

bs_4_2 <- reactive("ggplot(%oi, aes(x = %oi%ni)) + geom_bar() + ggtitle('%ni') + xlab('%ni')")

```

```{r, context = "server"}
output$plot_4_2 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_4_2(), 
      old_input = c(input$plot_dat_1, input$plot_var_1, input$plot_var_1),
      new_input = c(input$fill_1, input$plot_title_1, input$plot_xlab_1)
    )
   ))
})

output$code_4_2 <- renderText(
        makePrintText( 
          base_string = bs_4_2(), 
          old_input = c(input$plot_dat_1, input$plot_var_1, input$plot_var_1),
          new_input = c(input$fill_1, input$plot_title_1, input$plot_xlab_1)
    )
  )


```

```{r, echo = FALSE}
htmlOutput("code_4_2")

plotOutput("plot_4_2")
```

Now suppose we wish to compare two categorical variables.  One option is to simply make separate barplots for each category.  We do so by adding the option `facet_grid()` to our plot.  Try it below.

```{r, echo=FALSE}
textInput("plot_var_2", 
          "What variable would you like to use to separate the plots?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


bs_4_3 <- reactive("ggplot(%oi, aes(x = %oi%oi)) + geom_bar() + facet_grid(~%ni)")

```

```{r, context = "server"}
output$plot_4_3 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_4_3(), 
      old_input = c(input$plot_dat_1, input$plot_var_1, input$fill_1),
      new_input = c(input$plot_var_2)
    )
   ))
})

output$code_4_3 <- renderText(
        makePrintText( 
            base_string = bs_4_3(), 
            old_input = c(input$plot_dat_1, input$plot_var_1, input$fill_1, input$plot_var_1, input$plot_title_1, input$plot_xlab_1),
            new_input = c(input$plot_var_2)
    )
  )


```

```{r, echo = FALSE}
htmlOutput("code_4_3")

plotOutput("plot_4_3")
```

Another option is to make a **stacked bar chart**, which combines the two categorical variables on the bars themselves.  Note that these charts continue to obey the *area principle* - the area of the colored sections corresponds to the number of people in that category!


```{r, echo=FALSE}

textInput("plot_var_3_1", 
          "What variable would you like on the horizontal axis?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


textInput("plot_var_3_2", 
          "What variable would you like to include in the bar stacks?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


bs_4_4 <- reactive("ggplot(%oi, aes(x = %ni%ni)) + geom_bar()")

```

```{r, context = "server"}
output$plot_4_4 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_4_4(), 
      old_input = c("titanic"),
      new_input = c(input$plot_var_3_1, paste(", fill =", input$plot_var_3_2))
    )
   ))
})

output$code_4_4 <- renderText(
        makePrintText( 
          base_string = bs_4_4(), 
          old_input = c("titanic"),
          new_input = c(input$plot_var_3_1, paste(", fill =", input$plot_var_3_2))
    )
  )


```

```{r, echo = FALSE}
htmlOutput("code_4_4")

plotOutput("plot_4_4")
```


## Visualizing Quantitative Variables
Lastly, we will practice displaying quantitative variables.  Let us begin with a histogram.

(Hint: Go to [This Page](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf) to see lots of fun color options!)

```{r, echo=FALSE}
textInput("plot_var_5_1", 
          "What variable would you like to plot?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

sliderInput("plot_bins_5_1", 
          "How many bins should the histogram have?",
          value = 20,
          min = 1, 
          max = 100)

textInput("plot_color_5_1", 
          "What color should the histogram be?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


bs_5_1 <- reactive("ggplot(%oi) + geom_%oi(aes(x = %ni), bins = %ni, fill = '%ni')")
# 
htmlOutput("code_5_1")
# 
plotOutput("plot_5_1")
```


```{r, context="server"}
output$plot_5_1 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_5_1(), 
      old_input = c("titanic", "histogram"),
      new_input = c(input$plot_var_5_1, input$plot_bins_5_1, input$plot_color_5_1)
    )
   ))
})

output$code_5_1 <- renderText(
        makePrintText( 
          base_string = bs_5_1(), 
          old_input = c("titanic", "histogram"),
         new_input = c(input$plot_var_5_1, input$plot_bins_5_1, input$plot_color_5_1)
    )
  )

```

Recall from the previous section the options `xlab`, `ggtitle`, and `facet_grid`.

See if you can use these options, in the exact same way, to alter the code below so that it makes two beautiful histograms of the same quantitative variable across groups of a categorical variable.

```{r ex_5_1, exercise = TRUE}
ggplot(titanic) + geom_histogram(aes(), )
```


Finally, a better way to compare a quantitative and a categorical variable is to make **side-by-side boxplots**.


```{r, echo=FALSE}
textInput("plot_var_5_2", 
          "What QUANTITATIVE variable would you like to plot?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_var_5_3", 
          "What CATEGORICAL variable would you like to compare?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

bs_5_2 <- reactive("ggplot(%oi) + geom_%oi(aes(x = %ni, y = %ni, fill = %ni))")
# 
htmlOutput("code_5_2")
# 
plotOutput("plot_5_2")
```


```{r, context="server"}
output$plot_5_2 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_5_2(), 
      old_input = c("titanic", "boxplot"),
      new_input = c(input$plot_var_5_3, input$plot_var_5_2, input$plot_var_5_3)
    )
   ))
})

output$code_5_2 <- renderText(
        makePrintText( 
      base_string = bs_5_2(), 
      old_input = c("titanic", "boxplot"),
      new_input = c(input$plot_var_5_3, input$plot_var_5_2, input$plot_var_5_3)
    )
  )

```

## Finish up

Use the code box below to recreate your favorite plot.  Make sure it has a title and axis labels!


```{r ex_final, exercise = TRUE}
ggplot(titanic) + 
```

When you are happy with your plot, print this page (File > Save As) or (*Right Click* > Save As) and turn it in to the PreLab assignment on PolyLearn.

