---
title: "Lab Exercise: Quantitative Variables"
output: learnr::tutorial
tutorial:
  id: "Lab_Summaries"
  version: 1.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Lab_Exercise_QuantVars.html")){
  file.remove("Lab_Exercise_QuantVars.html")
}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)

knitr::opts_chunk$set(echo = FALSE)

tags$style(type="text/css",
         ".shiny-output-error { visibility: hidden; }",
         ".shiny-output-error:before { visibility: hidden; }"
)


source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/makeStrings.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')

titanic_full <- titanic %>% mutate(
   Family.Aboard = Siblings.Spouses.Aboard + Parents.Children.Aboard,
     Survived = factor(Survived,
                    levels = c(0,1),
                    labels = c("Died", "Lived")),
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)
### %ni = new input (red), %oi = old input (blue)
```


## Introduction

In this exercise, we will continue to explore the `titanic` dataset, which has already been loaded for you.  This time, we will think about **quantitative** variables - how to display them, how to summarize them, and how to answer research questions with plots and numerical summaries.

First, we will make the necessary adjustments to the categorical variables:

```{r, echo = TRUE}
titanic <- titanic %>% mutate(
  Survived = factor(Survived,
                    levels = c(0,1),
                    labels = c("Died", "Lived")),
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)
```

Then, remind yourself what the dataset looks like:

```{r, echo = TRUE}
titanic %>% head()
```


## Adding two variables

Suppose we are interested in combining information in the dataset.  For example, our `titanic` data contains two pieces of information:  The number of siblings or spouses travelling with a passenger, and the number of parents or children travelling with a passenger.

We may not be interested in this distinction; instead, we want to know how many family members overall are travelling with each passenger.  Let's choose to name this variable `Family.Aboard`.

To create this, we will `mutate()` our dataset to make our new variable. 


```{r add_vars}
textInput("sib_spouse",
          "What is the name of the variable containing the number of siblings or spouses travelling with each passenger?",
          value = "",
          width = '80%', placeholder = NULL)

textInput("parent_child",
          "What is the name of the variable containing the number of parents or children travelling with each passenger?",
          value = "",
          width = '80%', placeholder = NULL)


```

```{r, context="server"}

output$code_fam <- renderText(
        makePrintText(
          base_string = "%oi <- %oi %>% mutate( \n \t %oi = %ni + %ni \n )",
          old_input = c("titanic", "titanic", "Family.Aboard"),
          new_input = c(input$sib_spouse, input$parent_child)
        )
  )


```

```{r}
titanic <- titanic %>% mutate(
  Family.Aboard = Siblings.Spouses.Aboard + Parents.Children.Aboard
)
```

The code for this will look like:

```{r}
htmlOutput("code_fam")
```

Notice how we always use `<-` to update the dataset, i.e., to save the new calculations into our data.  Then, if we look at the data again, the new variable should be there:

```{r}
titanic %>% head()

```


## Histograms

To make other plots beyond the **barplots** that you already learned, you will use almost the same code.  We still use `ggplot()` in the same way, as well as options like `ggtitle()` and `xlab`.  The only difference is that instead of `geom_bar()`, we want to make a **histogram**, or `geom_histogram()`.  Try it below.  Pay careful attention to the new options that come with `geom_histogram()`

(Hint: Go to [This Page](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf) to see lots of fun color options!)

```{r, echo=FALSE}
textInput("plot_2_var", 
          "What QUANTITATIVE variable would you like to make a histogram of?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

sliderInput("plot_2_bins", 
          "How many bins should the histogram have?",
          value = 20,
          min = 1, 
          max = 100)

textInput("plot_2_color", 
          "What color should the histogram be?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_2_title", 
          "What title will you give your plot?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_2_xlab", 
          "What label should the x-axis have?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)



bs_2 <- reactive("ggplot(%oi, aes(x = %ni)) + geom_%oi(bins = %ni, %ni) + ggtitle('%ni') + xlab('%ni')")
# 
htmlOutput("code_2")
# 
plotOutput("plot_2")
```


```{r, context="server"}
output$plot_2 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_2(), 
      old_input = c("titanic_full", "histogram"),
      new_input = c(input$plot_2_var, input$plot_2_bins, paste0(", fill = '", input$plot_2_color, "'"), input$plot_2_title, input$plot_2_xlab)
    )
   ))
})

output$code_2 <- renderText(
        makePrintText( 
      base_string = bs_2(), 
      old_input = c("titanic", "histogram"),
      new_input = c(input$plot_2_var, input$plot_2_bins, paste0(", fill = '", input$plot_2_color, "'"), input$plot_2_title, input$plot_2_xlab)
    )
  )

```

### Densities

A histogram is our best guess about the shape of a **density**.  We can also choose to plot a smooth line over the histogram, to help us visualize what the density might be like.

```{r}

radioButtons("count_density",
             "The height of the bars should be:",
             choices = c("counts" = "", 
                         "densities" = "aes(y = ..density..), ")
)

radioButtons("add_density",
             "The plot should show an estimated density curve.",
             choices = c(
                "no" = "",
               "yes" = "+ geom_density(size = 2)"
             )
)
               


bs_3 <- reactive("ggplot(%oi, aes(x = %oi)) + geom_%oi(%nibins = %oi%oi) %ni+ ggtitle('%oi') + xlab('%oi')")
# 
htmlOutput("code_3")
# 
plotOutput("plot_3")
```

```{r, context="server"}
output$plot_3 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_3(), 
      old_input = c("titanic_full", input$plot_2_var, "histogram", input$plot_2_bins, paste0(", fill = '", input$plot_2_color, "'"), input$plot_2_title, input$plot_2_xlab),
      new_input = c(input$count_density, input$add_density)
    )
   ))
})

output$code_3 <- renderText(
        makePrintText( 
      base_string = bs_3(), 
      old_input = c("titanic", input$plot_2_var, "histogram", input$plot_2_bins, paste0(", fill = '", input$plot_2_color, "'"), input$plot_2_title, input$plot_2_xlab),
      new_input = c(input$count_density, input$add_density)
    )
  )

```

### Reading the histogram

Use the exercise above to with the variable `Age`. Think about the following questions:

* What number of bins one seems to give the most reasonable density estimate, in your opinion?

* How many modes does this distribution have?

* Is this distribution skewed?

* What are the **mean** and **median** age of Titanic passengers?


## Calculating summary statistics

The histogram let us "eyeball" reasonable guesses about the *mean* and *median* of our quantitative variables.  But of course, we can calculate these directly in **R**.

An easy method, and a good first step in any data analysis, is to use the function `summary()` to get some basic information about all the variables:

```{r}
titanic <- titanic_full
```

```{r, echo = TRUE}
titanic %>% summary()
```

### Counting categorical variables


```{r count}
textInput("var_0_1",
          "What is the first categorical variable you would you like to count?",
          value = "",
          width = '80%', placeholder = NULL)

textInput("var_0_2",
          "What is the second categorical variable you would you like to count?",
          value = "",
          width = '80%', placeholder = NULL)


bs_0 = reactive("%oi %>% count(%ni, %ni)")

```

```{r, context="server"}

output$code_0 <- renderText(
        makePrintText(
          base_string = bs_0(),
          old_input = c("titanic"),
          new_input = c(input$var_0_1, input$var_0_2)
        )
  )

output$output_0 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_0(),
          old_input = c("titanic"),
          new_input = c(input$var_0_1, input$var_0_2)
    )
   ))
})

```

```{r}
htmlOutput("code_0")
```

Result:

```{r}
verbatimTextOutput("output_0")
```

### Individual statistics

If we are only interested in a specific summary statistic, or if we want to calculate something that is not included in `summary()`, such as **standard deviation**, we can use the function `summarize_at()` to specify the variables and statistics to calculate.



```{r summarize}
textInput("var1",
          "Which QUANTITATIVE variable would you like to find information about?",
          value = "",
          width = '80%', placeholder = NULL)

checkboxGroupInput("funs1", 
          "What information do you want?",
          choices = c("Mean" = "mean",
                      "Median" = "median",
                      "Standard Deviation" = "sd",
                      "IQR" = "IQR",
                      "Sample Size" = "length")
)


bs_1_1 = reactive("%oi %>% summarize_at(vars(%ni), funs(%ni))")

```

```{r, context="server"}

output$code_1_1 <- renderText(
        makePrintText(
          base_string = bs_1_1(),
          old_input = c("titanic"),
          new_input = c(input$var1, toString(input$funs1))
        )
  )

output$output_1_1 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_1_1(), 
          old_input = c("titanic_full"),
          new_input = c(input$var1, toString(input$funs1))
    )
   ))
})

```

```{r}
htmlOutput("code_1_1")
```

Result:

```{r}
verbatimTextOutput("output_1_1")
```

### Summary statistics by Group

Some research questions will involve understanding how a quantitative variable relates to a categorical variable.  We may wish to calculate summary statistics separately for each category. We can achieve this with `group_by()`

```{r grooupby}
textInput("var2_quant",
          "Which QUANTITATIVE variable would you like to find information about?",
          value = "",
          width = '80%', placeholder = NULL)

textInput("var2_cat",
          "Which CATEGORICAL variable would you like to use to split groups?",
          value = "",
          width = '80%', placeholder = NULL)

checkboxGroupInput("funs2", 
          "What information do you want?",
          choices = c("Mean" = "mean",
                      "Median" = "median",
                      "Standard Deviation" = "sd",
                      "IQR" = "IQR",
                      "Sample Size" = "length")
)


bs_1_2 = reactive("%oi %>% group_by(%ni) %>% summarize_at(vars(%ni), funs(%ni))")

```

```{r, context="server"}

output$code_1_2 <- renderText(
        makePrintText(
          base_string = bs_1_2(),
          old_input = c("titanic"),
          new_input = c(input$var2_cat, input$var2_quant, toString(input$funs2))
        )
  )

output$output_1_2 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
        base_string = bs_1_2(),
        old_input = c("titanic"),
        new_input = c(input$var2_cat, input$var2_quant, toString(input$funs2))
    )
   ))
})

```

```{r}
htmlOutput("code_1_2")
```

Result:

```{r}
verbatimTextOutput("output_1_2")
```



### Check your work


```{r q_1.1}
textInput("a_1.1.1",
           "What is the standard deviation of the age of First Class passengers?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.1")

textInput("a_1.1.2",
           "What was the median age of Female passengers?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.2")

textInput("a_1.1.3",
           "What was the IQR for a ticket on the Titanic, based on this dataset?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1.3")
```

```{r, context = "server"}

output$check_1.1.1 <- renderText(
  if(input$a_1.1.1 == ""){
     ""
  }else if(round(as.numeric(input$a_1.1.1), 2) == round(sd(titanic[titanic$Pclass == 1, "Fare"]),2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.1.2 <- renderText(
   if(input$a_1.1.2 == ""){
     ""
  }else if(round(as.numeric(input$a_1.1.2),2) == round(median(titanic[titanic$Sex == "female", "Age"]),2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.1.3 <- renderText(
   if(input$a_1.1.3 == ""){
     ""
  }else if(round(as.numeric(input$a_1.1.3),2) == round(IQR(titanic$Fare, na.rm = TRUE), 2)){
      correct
  }else{
    incorrect
  }
)


```

## Comparing a Quantitative variable across categories

### Multiple Histograms

Recall from the previous lab the option `facet_grid`.  We can use this addition in the exact same way as we did for *barplots* to make multiple separate *histograms* across values a particular categorical variable.


```{r, echo=FALSE}
textInput("plot_4_var_quant", 
          "What QUANTITATIVE variable would you like to plot on a histogram?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_4_var_cat", 
          "What CATEGORICAL variable would you like to use to separate the histograms?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

bs_4 <- reactive("ggplot(%oi) + geom_histogram(aes(x = %ni)) + %oi(~%ni)")
# 
htmlOutput("code_4")
# 
plotOutput("plot_4")
```


```{r, context="server"}
output$plot_4 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_4(), 
      old_input = c("titanic_full", "facet_grid"),
      new_input = c(input$plot_4_var_quant, input$plot_4_var_cat)
    )
   ))
})

output$code_4 <- renderText(
        makePrintText( 
      base_string = bs_4(), 
      old_input = c("titanic", "facet_grid"),
      new_input = c(input$plot_4_var_quant, input$plot_4_var_cat)
    )
  )

```

### Boxplots

A better way to compare a quantitative and a categorical variable is usually to make **side-by-side boxplots**.


```{r, echo=FALSE}
textInput("plot_5_var_quant", 
          "What QUANTITATIVE variable would you like to plot?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_5_var_cat", 
          "What CATEGORICAL variable would you like to compare?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

bs_5 <- reactive("ggplot(%oi) + geom_%oi(aes(x = %ni, y = %ni, fill = %ni))")
# 
htmlOutput("code_5")
# 
plotOutput("plot_5")
```


```{r, context="server"}
output$plot_5 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_5(), 
      old_input = c("titanic_full", "boxplot"),
      new_input = c(input$plot_5_var_cat, input$plot_5_var_quant, input$plot_5_var_cat)
    )
   ))
})

output$code_5 <- renderText(
        makePrintText( 
      base_string = bs_5(), 
      old_input = c("titanic", "boxplot"),
      new_input = c(input$plot_5_var_cat, input$plot_5_var_quant, input$plot_5_var_cat)
    )
  )

```