---
title: "Lab Exercise: Binomial Random Variables"
output: learnr::tutorial
tutorial:
  id: "Lab_Binomial"
  version: 1.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)

knitr::opts_chunk$set(echo = FALSE)


source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/makeStrings.R")
source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/plotDensities.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

```


## Binomial Probabilities in R

We will be using **R** to calculate information about Binomial random variables.  The first important function is `dbinom()`, which finds the *probability* of some number of successes.


```{r factor_var}
textInput("n",
          "How many trials do you have?",
          value = "0",
          width = '80%', placeholder = NULL)


textInput("p",
          "What is the probability of success at each trial?",
          value = "0",
          width = '80%', placeholder = NULL)

textInput("k",
          "We are interested in the probability of how many successes?",
          value = "0",
          width = '80%', placeholder = NULL)

bs_1_1 = reactive("dbinom(%ni, size = %ni, prob = %ni)")

```

```{r, context="server"}

output$code_1_1 <- renderText(
        makePrintText(
          base_string = bs_1_1(),
          old_input = c(),
          new_input = c(input$k, input$n, input$p)
        )
  )

output$output_1_1 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_1_1(), 
      old_input = c(),
      new_input = c(input$k, input$n, input$p)
    )
   ))
})

output$binom_var_1 <- renderText(
        makePrintText(
          base_string = "X ~ Binomial(%ni, %ni) \n P(X = %ni)",
          old_input = c(),
          new_input = c(input$n, input$p, input$k)
        )
  )


output$distn_1 <- renderPlot(
  illustrate_binom(input$n, input$p, x = input$k, equal_to = TRUE)
)

```

Code:
```{r}
htmlOutput("code_1_1")
```

Result:

```{r}
verbatimTextOutput("output_1_1")
```

Random variable and probability:
```{r}
htmlOutput("binom_var_1")
```

Illustration:
```{r}
plotOutput("distn_1")
```

It is not very common that we are interested in the probability of some exact number of successes.  More often, we want to quantify evidence by finding the probability of some number **or more**.  For this, we will use *pbinom()*.  By default, `pbinom()` finds the probability of getting **less than or equal to** the given number.  Try it below.

```{r pbinom}
textInput("n2",
          "How many tries do you have?",
          value = "0",
          width = '80%', placeholder = NULL)


textInput("p2",
          "What is the probability of success at each try?",
          value = "0",
          width = '80%', placeholder = NULL)

textInput("k2",
          "We are interested in the probability of less than or equal to how many successes?",
          value = "0",
          width = '80%', placeholder = NULL)

bs_1_2 = reactive("pbinom(%ni, size = %ni, prob = %ni)")

```

```{r, context="server"}

output$code_1_2 <- renderText(
        makePrintText(
          base_string = bs_1_2(),
          old_input = c(),
          new_input = c(input$k2, input$n2, input$p2)
        )
  )

output$output_1_2 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_1_2(), 
      old_input = c(),
      new_input = c(input$k2, input$n2, input$p2)
    )
   ))
})

```

```{r}
htmlOutput("code_1_2")
```

Result:

```{r}
verbatimTextOutput("output_1_2")
```

Finally, sometimes we want to ask probability questions "in reverse".  For example, we might ask "If I flip 100 coins, I can be 95% sure that I will get at least how many tails?".  We could answer this via guess-and-check with `pbinom()`, but fortunately, there is a function for it in **R** called `qbinom()`.

It is not very common that we are interested in the probability of some exact number of successes.  More often, we want to quantify evidence by finding the probability of some number **or more**.  For this, we will use *pbinom()*.  By default, `pbinom()` finds the probability of getting **less than or equal to** the given number.  Try it below.

```{r qbinom}
textInput("n3",
          "How many tries do you have?",
          value = "0",
          width = '80%', placeholder = NULL)


textInput("p3",
          "What is the probability of success at each try?",
          value = "0",
          width = '80%', placeholder = NULL)

textInput("q",
          "We want to find the number of successes that has a probability of ____ below it.",
          value = "0",
          width = '80%', placeholder = NULL)

bs_1_3 = reactive("qbinom(%ni, size = %ni, prob = %ni)")

```

```{r, context="server"}

output$code_1_3 <- renderText(
        makePrintText(
          base_string = bs_1_3(),
          old_input = c(),
          new_input = c(input$q, input$n3, input$p3)
        )
  )

output$output_1_3 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_1_3(), 
      old_input = c(),
      new_input = c(input$q, input$n3, input$p3)
    )
   ))
})

```

```{r}
htmlOutput("code_1_3")
```

Result:

```{r}
verbatimTextOutput("output_1_3")
```

### Practice finding probabilities

Use the following code box to run any code you need for the following questions.

```{r ex_1.1, exercise = TRUE}
```

```{r q_1}
textInput("a_1.1",
           "I flip 100 fair coins.  What is the probability that I get exactly 50 tails?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1")

textInput("a_1.2",
           "I flip 100 fair coins.  What is the probability that I get less than 45 tails?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.2")

textInput("a_1.3",
           "About 10% of all humans are left-handed.  If I randomly choose 43 humans, what is the probability that at least 8 of them are left handed?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.3")

textInput("a_1.4",
           "About 10% of all humans are left-handed.  In a class of 20 people, I can be 80% sure that less than how many people are left-handed?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.4")
```

```{r, context = "server"}

output$check_1.1 <- renderText(
  if(input$a_1.1 == ""){
     ""
  }else if(round(input$a_1.1,2) == round(dbinom(50, 100, .5), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.2 <- renderText(
  if(input$a_1.2 == ""){
    ""
  }else if(round(input$a_1.2,2) == round(pbinom(44, 100, .5), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.3 <- renderText(
  if(input$a_1.3 == ""){
     ""
  }else if(round(input$a_1.3,2) == round(1-pbinom(7, 43, .1), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.4 <- renderText(
  if(input$a_1.4 == ""){
     ""
  }else if(round(input$a_1.4,2) == round(qbinom(.8, 20, .1), 2)){
      correct
  }else{
    incorrect
  }
)


```


### Practice predicting code

A good habit to get into when coding is to "translate" your code to English in your head, and try to predict what the output will be.  This is how to figure out if you are truly asking the computer to calculate what you hoped.  (Remember - computers will almost always output *something*!  It is your responsibility as the scientist to make sure you have asked the right question.)

For the following, try to guess what the output would be.  Do **not** run the code; instead, try to read it as a probability question and think through the answer.


*Give answers to 2 decimals, and make sure to put a 0 in front, e.g., 0.67 not .6*


```{r q_2}
textInput("a_2.1",
           "dbinom(1, size = 1, prob = 0.7)",
           value = "0",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.1")

textInput("a_2.2",
           "pbinom(60, size = 60, prob = 0.13)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.2")

textInput("a_2.4",
           "qbinom(1, size = 60, prob = 0.13)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.4")


textInput("a_2.3",
           "pbinom(50, size = 100, prob = 0.5)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.3")

textInput("a_2.5",
           "qbinom(0.5, size = 100, prob = 0.5)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.5")
```

```{r, context = "server"}

output$check_2.1 <- renderText(
  if(input$a_2.1 == ""){
     ""
  }else if(input$a_2.1 == 0.7){
      correct
  }else{
    incorrect
  }
)

output$check_2.2 <- renderText(
  if(input$a_2.2 == ""){
     ""
  }else if(input$a_2.2 == 1){
      correct
  }else{
    incorrect
  }
)

output$check_2.3 <- renderText(
  if(input$a_2.3 == ""){
     ""
  }else if(input$a_2.3 == 0.5){
      correct
  }else{
    incorrect
  }
)

output$check_2.4 <- renderText(
  if(input$a_2.4 == ""){
     ""
  }else if(input$a_2.4 == 60){
      correct
  }else{
    incorrect
  }
)

output$check_2.5 <- renderText(
  if(input$a_2.5 == ""){
     ""
  }else if(input$a_2.5 == 50){
      correct
  }else{
    incorrect
  }
)


```