---
title: "Lab 7 "
author: "YOUR NAME HERE"
date: "November 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
### YOU CAN IGNORE EVERYTHING IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)

source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/packageLoaders2.R")

require(ggplot2)
require(dplyr)
require(tidyr)
require(reshape2)
require(fmsb)

babies = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/babies_full.csv')


```


The dataset `babies` contains baby names in California.   

Use this dataset to discover something interesting about the counts of babies with certain names.  You can consider any question you want - your own name, your family and friends, or something else!

Your discussion should include:
- A general overview and summary of the dataset.
- A statement of at least one clear research question that is answered with ANOVA
- A plot that addresses your research question
- An ANOVA test, Tukey Test, and discussion
- A statement of at least one clear research question that is answered with regression.
- A regression calculation, plot, and discussion.


For extra credit up to 10 points, also include a third research question and complete analysis that uses other methods from the class.

```{r}
## A code chunk
```



