---
title: "Lab Exercise: Chi-square and risk ratios"
output: learnr::tutorial
tutorial:
  id: "Lab_chisq"
  version: 2.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Lab_Exercise_Chi_Square.html")){
  file.remove("Lab_Exercise_Chi_Square.html")
}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)
require(tibble)

if(!require(fmsb)){
  install.packages("fmsb", repos = "https://cloud.r-project.org")
}

require(fmsb)

knitr::opts_chunk$set(echo = FALSE)

tags$style(type="text/css",
         ".shiny-output-error { visibility: hidden; }",
         ".shiny-output-error:before { visibility: hidden; }"
)


source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/makeStrings.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')

titanic <- titanic %>% mutate(
   Family.Aboard = Siblings.Spouses.Aboard + Parents.Children.Aboard,
     Survived = factor(Survived,
                    levels = c(0,1),
                    labels = c("Died", "Lived")),
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)
### %ni = new input (red), %oi = old input (blue)
```


## Introduction

By now, you no doubt had the `titanic` dataset nearly memorized!  But just in case, here is some information for your reference in this exercise.

```{r, echo = TRUE}
titanic %>% head()
titanic %>% summary()
```


## Two-Way Tables

Let's begin by practicing creating two-way tables, and calculating their important summary distributions, using **R**.  In the `Categorical Variables` Lab Exercise, you learned to use `count()` to find the number of observations in each category of one or more variables.  We will now use `spread()` to reformat this information into a two-way table.  Note that we `spread()` the information horizontally, by the **explanatory** variable.


```{r, echo=FALSE}
textInput("var_1", 
          "What categorical variable will be your EXPLANATORY variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("var_2", 
          "What categorical variable will be your RESPONSE variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

bs_1 <- reactive("%oi %>% count(%ni, %ni) %>% spread(%ni, n)")

```

```{r, context="server"}

output$code_1 <- renderText(
        makePrintText(
          base_string = bs_1(),
          old_input = c("titanic"),
          new_input = c(input$var_1, input$var_2, input$var_1)
        )
  )

output$output_1 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_1(),
          old_input = c("titanic"),
          new_input = c(input$var_1, input$var_2, input$var_1)
    )
   ))
})

```

Code:
```{r}
htmlOutput("code_1")
```

Result from code:
```{r}
verbatimTextOutput("output_1")
```

### Row and column sums

Now that we have found the two-way table of counts, we are probably interested in filling in the totals as well.  **R** makes this easy with `sum()`, `rowSums()` and `colSums()`.  Try it below.

Note: we first use `select()` to let **R** know that we want to ignore the row labels, and just find the sum of the counts.

```{r}

radioButtons("sum_type",
           "We would like to find the",
           choices = c(
             "row totals" = "rowSums",
              "column totals" = "colSums",
              "overall total" = "sum")
)

bs_2 <- reactive("tw <- %oi %>% count(%ni, %ni) %>% spread(%ni, n)
                 tw %>% select(-%ni) %>% %ni()")

```


```{r, context="server"}

output$code_2 <- renderText(
        makePrintText(
          base_string = bs_2(),
          old_input = c("titanic"),
          new_input = c(input$var_1, input$var_2, input$var_1, input$var_2, input$sum_type)
        )
  )

output$output_2 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_2(),
          old_input = c("titanic"),
          new_input = c(input$var_1, input$var_2, input$var_1, input$var_2, input$sum_type)
    )
   ))
})

```

Code:
```{r}
htmlOutput("code_2")
```

Result from code:
```{r}
verbatimTextOutput("output_2")
```


## Joint, marginal, and conditional distributions

Let's now look at using `Sex` as the explanatory variable, and `Survived` as the response.  First, instead of simply calculating our counts and totals, we will have **R** save this information.

```{r, echo = TRUE}
counts <- titanic %>% count(Sex, Survived) %>% spread(Sex, n) %>% select(-Survived)

column_totals <- counts %>% colSums()
row_totals <- counts %>% rowSums()
overall_total <- counts %>% sum()

```

To find the joint distribution, we are interested in the original counts in each subcategory, divided by the overall total number of people.

```{r, echo = TRUE}
counts/overall_total
```

To find the marginal distribution of `Sex`, we are interested in the total counts for each category of the explanatory variable (columns) divided by the total number of people.

```{r, echo = TRUE}
column_totals/overall_total
```

To find the marginal distribution of `Survived`, we are interested in the total counts for each category of the response variable (rows) divided by the total number of people.

```{r, echo = TRUE}
row_totals/overall_total
```


Finally, for the conditional distribution, we are interested in focusing on each category of the explanatory variable - `male` or `female`.  Thus, we want to see the counts in each quadrant out of the total number of women or men (column totals).

```{r, echo = TRUE}
t(t(counts)/column_totals)
```


### Practice

Use the exercises above, and the code box below, to answer the following questions.

```{r ex_1, exercise = TRUE}

```

```{r q_1}


textInput("a_1.1",
           "How many Second Class passengers survived the sinking?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1")



textInput("a_1.2",
           "What proportion of passengers travelled in Second Class?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.2")

textInput("a_1.3",
           "What proportion of passengers survived the sinking?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.3")

textInput("a_1.4",
           "What proportion of Second Class passengers survived the sinking?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.4")


```

```{r, context = "server"}

output$check_1.1 <- renderText(
  if(input$a_1.1 == ""){
     ""
  }else if(as.numeric(input$a_1.1) == sum(titanic$Survived == "Lived" & titanic$Passenger.Class == "Second")){
      correct
  }else{
    incorrect
  }
)
    
output$check_1.2 <- renderText(
  if(input$a_1.2 == ""){
     ""
  }else if(round(as.numeric(input$a_1.2), 2) == round(sum(titanic$Passenger.Class == "Second")/nrow(titanic), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.3 <- renderText(
  if(input$a_1.3 == ""){
     ""
  }else if(round(as.numeric(input$a_1.3), 2) == round(sum(titanic$Survived == "Lived")/nrow(titanic), 2)){
      correct
  }else{
    incorrect
  }
)
output$check_1.4 <- renderText(
  if(input$a_1.4 == ""){
     ""
  }else if(round(as.numeric(input$a_1.4), 2) == round(sum(titanic$Passenger.Class == "Second" & titanic$Survived == "Lived")/sum(titanic$Passenger.Class == "Second"), 2)){
      correct
  }else{
    incorrect
  }
)
```


## Chi-Squared Tests

Now explore using `chisq.test()` to perform hypothesis tests on two-way tables.

```{r, echo=FALSE}
textInput("var3_1", 
          "What categorical variable will be your EXPLANATORY variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("var3_2", 
          "What categorical variable will be your RESPONSE variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

bs_3 <- reactive("tw <- %oi %>% count(%ni, %ni) %>% spread(%ni, n) %>% select(-%ni)
                 chisq.test(tw)")

```

```{r, context="server"}

output$code_3 <- renderText(
        makePrintText(
          base_string = bs_3(),
          old_input = c("titanic"),
          new_input = c(input$var3_1, input$var3_2, input$var3_1, input$var3_2)
        )
  )

output$output_3 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_3(),
          old_input = c("titanic"),
          new_input = c(input$var3_1, input$var3_2, input$var3_1, input$var3_2)
    )
   ))
})

output$hypos_3 <- renderText(
        makePrintText(
          base_string = "H0: There is no relationship between %ni and %ni.\n  Ha: There is a relationship between %ni and %ni.",
          old_input = c(),
          new_input = c(input$var3_1, input$var3_2, input$var3_1, input$var3_2)
        )
  )

```

Hypotheses:
```{r}
htmlOutput("hypos_3")
```

Code:
```{r}
htmlOutput("code_3")
```

Result from code:
```{r}
verbatimTextOutput("output_3")
```

### Practice

Test whether a person's travel class relates to his or her survival.

```{r ex_2, exercise = TRUE}

```

```{r q_2}


textInput("a_2.1",
           "What was the test statistic?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.1")



textInput("a_2.2",
           "What was the p-value of this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.2")


radioButtons("a_2.3",
           "At the 0.05 level, we conclude that",
           choices = c(
             "There is not enough evidence that passenger class mattered to survival." = "no1",
              "There is strong evidence that passenger class mattered to survival." = "yes",
              "We cannot tell for sure from this test." = "no2")
)

htmlOutput("check_2.3")


```

```{r, context = "server"}

chisq_test <- reactive(titanic %>% count(Passenger.Class, Survived) %>% spread(Survived, n) %>% select(-Passenger.Class) %>% chisq.test())

output$check_2.1 <- renderText(
  if(input$a_2.1 == ""){
     ""
  }else if(round(as.numeric(input$a_2.1),2) == round(chisq_test()$statistic, 2)){
      correct
  }else{
    incorrect
  }
)


output$check_2.2 <- renderText(
  if(input$a_2.2 == ""){
     ""
  }else if(round(as.numeric(input$a_2.2),2) == round(chisq_test()$p.value, 2)){
      correct
  }else{
    incorrect
  }
)

output$check_2.3 <- renderText(
  if(input$a_2.3 == ""){
     ""
  }else if(input$a_2.3 == "yes"){
      correct
  }else{
    incorrect
  }
)

```


## Relative Risk

Now explore using `riskratio()` to perform a hypothesis test about gender and survival on the titanic.  You may need to use the code box below to find some of your inputs.

Note:  By default, this function will calculate p-values for the alternate hypothesis "The risk ratio is greater than one."

```{r ex_4_1, exercise = TRUE}

```

```{r, echo=FALSE}
textInput("count4_1_1", 
          "How many women are there in this dataset?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("count4_1_2", 
          "How many women in this dataset did NOT survive the sinking?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("count4_2_1", 
          "How many men are there in this dataset?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("count4_2_2", 
          "How many men in this dataset did NOT survive the sinking?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)



bs_4 <- reactive("riskratio(%ni, %ni, %ni, %ni)")

```

```{r, context="server"}

output$code_4 <- renderText(
        makePrintText(
          base_string = bs_4(),
          old_input = c(),
          new_input = c(input$count4_1_2, input$count4_2_2, input$count4_1_1, input$count4_2_1)
        )
  )

output$output_4 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_4(),
          old_input = c(),
          new_input = c(input$count4_1_2, input$count4_2_2, input$count4_1_1, input$count4_2_1)
    )
   ))
})


```

Code:
```{r}
htmlOutput("code_4")
```

Result from code:
```{r}
verbatimTextOutput("output_4")
```


### Practice

Suppose you want to test the hypothesis that men were more likely to die on the titanic than women.  Use the code box below to answer the questions.

```{r ex_4, exercise = TRUE}

```

```{r q_4}


textInput("a_4.1",
           "What was the test statistic?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_4.1")



textInput("a_4.2",
           "What was the p-value of this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_4.2")


radioButtons("a_4.3",
           "At the 0.05 level, we conclude that",
           choices = c(
             "Women were more likely to die on the Titanic than men." = "no1",
              "Men were more likely to die on the Titanic than women." = "yes",
             "Women and men were not equally likely to die on the Titanic." = "no2",
              "There was not strong evidence of a difference in genders and survival." = "no3")
)

htmlOutput("check_4.3")


```

```{r, context = "server"}

rr <- reactive(riskratio(464, 81, 464+109, 81+233))

output$check_4.1 <- renderText(
  if(input$a_4.1 == ""){
     ""
  }else if(round(as.numeric(input$a_2.1),2) == round(rr()$estimate, 2)){
      correct
  }else{
    incorrect
  }
)


output$check_4.2 <- renderText(
  if(input$a_4.2 == ""){
     ""
  }else if(round(as.numeric(input$a_2.2),2) == round(rr()$p.value, 2)){
      correct
  }else{
    incorrect
  }
)

output$check_4.3 <- renderText(
  if(input$a_4.3 == ""){
     ""
  }else if(input$a_4.3 == "yes"){
      correct
  }else{
    incorrect
  }
)

```