---
title: "Lab Exercise: Linear Regression"
output: learnr::tutorial
tutorial:
  id: "Lab_regression"
  version: 2.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Lab_Exercise_Linear_Regression.html")){
  file.remove("Lab_Exercise_Linear_Regression.html")
}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)
require(tibble)
require(fmsb)

knitr::opts_chunk$set(echo = FALSE)

tags$style(type="text/css",
         ".shiny-output-error { visibility: hidden; }",
         ".shiny-output-error:before { visibility: hidden; }"
)


source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/makeStrings.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

babies = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/babies_sub.csv')
kellys <- babies %>% filter(Name == "Kelly", Gender == "F")

# titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')
# 
# titanic <- titanic %>% mutate(
#    Family.Aboard = Siblings.Spouses.Aboard + Parents.Children.Aboard,
#      Survived = factor(Survived,
#                     levels = c(0,1),
#                     labels = c("Died", "Lived")),
#   Passenger.Class = factor(Pclass, 
#                            levels = c(1,2,3), 
#                            labels = c("First", "Second", "Third"))
# )
### %ni = new input (red), %oi = old input (blue)
```

## Baby Names in the U.S.

Our dataset in this lab concerns baby names in a few different states.  

The loaded dataset, `babies`, has information about baby names in the United States from 1940-2016.  For each year, and for each name with at least 50 recorded babies born, we are given the counts of how many babies had that name.  In this dataset, we include only California, and only a few baby names.

Take a look at the dataset, and familiarize yourself with the variables.

```{r ex_1.1, exercise = TRUE}
```


## Scatterplots

Let's take a look at how the name "Kelly" has changed over time, for female babies only.  First, we will make a new dataset called "kellys", that contains only counts for female Kellys.

```{r, echo = TRUE}
kellys <- babies %>% filter(Name == "Kelly", Gender == "F")
```


Use the following code to plot the counts of babies named Kelly (in California) for each year.  Notice the new addition to `ggplot()` of `+ geom_point()`


```{r, echo=FALSE}
textInput("plot_var_2_1", 
          "What variable would you like to plot on the x-axis?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_var_2_2", 
          "What variable would you like to plot on the y-axis?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

radioButtons("plot_type", 
             "Would you like a scatterplot or a line?",
                   choices = c("Scatterplot" = "point",
                               "Line" = "line"
                              )
              )

bs_2 <- reactive("ggplot(%oi, aes(x = %ni, y = %ni)) + geom_%ni()")
# 
htmlOutput("code_2")
# 
plotOutput("plot_2")
```


```{r, context="server"}
output$plot_2 <- renderPlot({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_2(), 
      old_input = c("kellys"),
      new_input = c(input$plot_var_2_1, input$plot_var_2_2, input$plot_type)
    )
   ))
})

output$code_2 <- renderText(
        makePrintText( 
      base_string = bs_2(), 
      old_input = c("kellys"),
      new_input = c(input$plot_var_2_1, input$plot_var_2_2, input$plot_type)
    )
  )

```

**Think about it:**  Why did the name "Kelly" suddenly become popular for female babies around 1960?  I have heard a few theories... can you find me an interesting one?  

## Regression

It appears that my name is getting steadily less popular since 1980.  However, we would like to know if this trend is significant.  First, let's adjust our dataset to narrow it down to only 1980 and beyond.

```{r, echo = TRUE}
kellys_80 <- babies %>% filter(Name == "Kelly", Gender == "F", Year > 1980)
```

Use the code below to run a **linear regression** on the counts of Kellys over time.  For this, we use the **R** command `lm()`, which stands for "linear model". 



```{r, echo=FALSE}
textInput("plot_var_3_1", 
          "What is the explanatory variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_var_3_2", 
          "What is the response variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

bs_3 <- reactive("summary(%oi(%ni ~ %ni, data = %oi))")
# 
htmlOutput("code_3")
# 
verbatimTextOutput("plot_3")
```


```{r, context="server"}
output$plot_3 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_3(), 
      old_input = c("lm", "kellys_80"),
      new_input = c(input$plot_var_3_2, input$plot_var_3_1)
    )
   ))
})

output$code_3 <- renderText(
        makePrintText( 
      base_string = bs_3(), 
      old_input = c("lm", "kellys_80"),
      new_input = c(input$plot_var_3_2, input$plot_var_3_1)
    )
  )

```


```{r q_3}
textInput("a_3",
           "What is the test statistic for reduction in variance due to this regression line?",
           value = "0",
           width = '80%', placeholder = NULL)

htmlOutput("check_3")

```

```{r, context = "server"}

output$check_3 <- renderText(
  if(round(as.numeric(input$a_3), 2) == 1565){
      correct
  }else{
    incorrect
  }
)
```


```{r q_31}
textInput("a_31",
           "What is the test statistic for the slope of this regression line?",
           value = "0",
           width = '80%', placeholder = NULL)

htmlOutput("check_31")

```

```{r, context = "server"}

output$check_31 <- renderText(
  if(round(as.numeric(input$a_31), 2) == -39.55){
      correct
  }else{
    incorrect
  }
)
```

```{r q_4}
textInput("a_4",
           "What was the p-value for this regression line?",
           value = "1",
           width = '80%', placeholder = NULL)

htmlOutput("check_4")

```

```{r, context = "server"}

output$check_4 <- renderText(
  if(as.numeric(input$a_4) == 0){
      correct
  }else{
    incorrect
  }
)
```



```{r q_6}
textInput("a_6",
           "What percent of variance in the number of Kellys born in a year is explained by a change over time?",
           value = "0",
           width = '80%', placeholder = NULL)

htmlOutput("check_6")

```

```{r, context = "server"}

output$check_6 <- renderText(
  if(round(as.numeric(input$a_6), 2) == 97.87){
      correct
  }else{
    incorrect
  }
)
```


```{r q_5}
textInput("a_5",
           "What is the correlation between number of Kellys and time?",
           value = "0",
           width = '80%', placeholder = NULL)

htmlOutput("check_5")

```

```{r, context = "server"}

output$check_5 <- renderText(
  if(round(as.numeric(input$a_5), 2) == -0.99){
      correct
  }else{
    incorrect
  }
)
```

## Checking residuals

The last thing we need to do is make sure the *linear* regression is a good choice.  Use the code below to plot the regression line.

```{r, echo=FALSE}
textInput("plot_var_4_1", 
          "What is your explanatory variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("plot_var_4_2", 
          "What is your response variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

bs_4 <- reactive("my_lm <- lm(%ni ~ %ni, data = %oi)")
# 
htmlOutput("code_4")
```


```{r, context="server"}

output$code_4 <- renderText(
        makePrintText( 
      base_string = bs_4(), 
      old_input = c("kellys_80", "point"),
      new_input = c(input$plot_var_4_1, input$plot_var_4_2, input$reg_line)
    )
  )

```

Here is a plot of the residuals. In the future, you can simply copy this code, once you have made your own `my_lm`.


```{r, echo = FALSE}
my_lm <- lm(Count ~ Year, data = kellys_80)
```

```{r, echo = TRUE}

ggplot(my_lm) + 
  geom_point(aes(x=.fitted, y=.resid)) +
  xlab("Fitted values") + ylab("Residuals") +
  geom_hline(yintercept = 0, col = "red")

```

```{r, echo=FALSE}
textAreaInput("discuss_1", 
          "Do you see any issue with the regression, based on the residuals?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

```
