---
title: "Lab Exercise: Binomial Random Variables"
output: learnr::tutorial
tutorial:
  id: "Lab_Binomial"
  version: 1.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)

knitr::opts_chunk$set(echo = FALSE)


source("https://raw.githubusercontent.com/kbodwin/Introductory-Statistics-Labs/master/Scripts/makeStrings.R")
source("https://raw.githubusercontent.com/kbodwin/Introductory-Statistics-Labs/master/Scripts/plotDensities.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

```


## Binomial Probabilities in R

We will be using **R** to calculate information about Binomial random variables.  The first important function is `dbinom()`, which finds the *probability* of some number of successes.


```{r factor_var}
textInput("n",
          "How many trials do you have?",
          value = "0",
          width = '80%', placeholder = NULL)


textInput("p",
          "What is the probability of success at each trial?",
          value = "0",
          width = '80%', placeholder = NULL)

textInput("k",
          "We are interested in the probability of how many successes?",
          value = "0",
          width = '80%', placeholder = NULL)

bs_1_1 = reactive("dbinom(%ni, size = %ni, prob = %ni)")

```

```{r, context="server"}

output$code_1_1 <- renderText(
        makePrintText(
          base_string = bs_1_1(),
          old_input = c(),
          new_input = c(input$k, input$n, input$p)
        )
  )

output$output_1_1 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_1_1(), 
      old_input = c(),
      new_input = c(input$k, input$n, input$p)
    )
   ))
})

output$binom_var_1 <- renderText(
        makePrintText(
          base_string = "X ~ Binomial(%ni, %ni) \n P(X = %ni) = ?",
          old_input = c(),
          new_input = c(input$n, input$p, input$k)
        )
  )


output$distn_1 <- renderPlot(
  illustrate_binom(input$n, input$p, x = input$k, equal_to = TRUE)
)

```

Random variable and probability statement:
```{r}
htmlOutput("binom_var_1")
```

Code:
```{r}
htmlOutput("code_1_1")
```

Result:
```{r}
verbatimTextOutput("output_1_1")
```


Illustration:
```{r}
plotOutput("distn_1")
```

### Your turn

In the box below, you may run any line of **R** code.  Use the box, and/or the exercise above, to answer the following questions.


```{r ex_1.1, exercise = TRUE}

```

```{r q_1}
textInput("a_1.1",
           "I flip 30 fair coins.  What is the probability that I get exactly 15 tails?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1")

textInput("a_1.2",
           "16% of all Starburst candies in a pack are cherry flavored.  What is the probability that in a pack of 12, I get exactly 4 cherry Starbursts?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.2")

textInput("a_1.3",
           "In an upcoming election, 40% of voters prefer Candidate A, and 60% prefer Candidate B.  If only 16 people vote, what is the probability that the vote is NOT tied?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.3")

```

```{r, context = "server"}

output$check_1.1 <- renderText(
  if(input$a_1.1 == ""){
     ""
  }else if(round(input$a_1.1,2) == round(dbinom(15, 30, .5), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.2 <- renderText(
  if(input$a_1.2 == ""){
     ""
  }else if(round(input$a_1.2,2) == round(dbinom(4, 12, 0.16), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.3 <- renderText(
  if(input$a_1.3 == ""){
     ""
  }else if(round(input$a_1.3,2) == round(1- dbinom(8, 16, 0.4), 2)){
      correct
  }else{
    incorrect
  }
)
```

## More complicated probabilities

It is not very common that we are interested in the probability of some exact number of successes.  More often, we want to quantify **evidence** by finding the probability of some number **or more** of successes.  For this, we will use *pbinom()*.  By default, `pbinom()` finds the probability of getting **less than or equal to** the given number.  Try it below.

```{r pbinom}
textInput("n2",
          "How many trials do you have?",
          value = "0",
          width = '80%', placeholder = NULL)


textInput("p2",
          "What is the probability of success at each trial?",
          value = "0",
          width = '80%', placeholder = NULL)

textInput("k2",
          "We are interested in the probability of less than or equal to how many successes?",
          value = "0",
          width = '80%', placeholder = NULL)

bs_1_2 = reactive("pbinom(%ni, size = %ni, prob = %ni)")

```

```{r, context="server"}

output$code_2 <- renderText(
        makePrintText(
          base_string = bs_1_2(),
          old_input = c(),
          new_input = c(input$k2, input$n2, input$p2)
        )
  )

output$output_2 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_1_2(), 
      old_input = c(),
      new_input = c(input$k2, input$n2, input$p2)
    )
   ))
})

output$binom_var_2 <- renderText(
        makePrintText(
          base_string = "X ~ Binomial(%ni, %ni) \n P(X ≤ %ni) = ?",
          old_input = c(),
          new_input = c(input$n2, input$p2, input$k2)
        )
  )


output$distn_2 <- renderPlot(
  illustrate_binom(input$n2, input$p2, x = input$k2)
)

```


Random variable and probability statement:
```{r}
htmlOutput("binom_var_2")
```

Code:
```{r}
htmlOutput("code_2")
```

Result:
```{r}
verbatimTextOutput("output_2")
```


Illustration:
```{r}
plotOutput("distn_2")
```

### Your turn

```{r ex_2, exercise = TRUE}

```

```{r q_2}
textInput("a_2.1",
           "I flip 30 fair coins.  What is the probability that I get less than half tails?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.1")

textInput("a_2.2",
           "16% of all Starburst candies in a pack are cherry flavored.  What is the probability that in a pack of 12, I get more than 4 cherry Starbursts?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.2")

textInput("a_2.3",
           "About 10% of all humans are left-handed.  If I randomly choose 43 humans, what is the probability that at least 8 of them are left handed?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.3")

```

```{r, context = "server"}

output$check_2.1 <- renderText(
  if(input$a_2.1 == ""){
     ""
  }else if(round(input$a_2.1,2) == round(pbinom(14, 30, .5), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_2.2 <- renderText(
  if(input$a_2.2 == ""){
     ""
  }else if(round(input$a_2.2,2) == round(1-pbinom(4, 12, 0.16), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_2.3 <- renderText(
  if(input$a_2.3 == ""){
     ""
  }else if(round(input$a_2.3,2) == round(1- pbinom(7, 43, 0.1), 2)){
      correct
  }else{
    incorrect
  }
)
```

## Finding values from probabilities

Finally, sometimes we want to ask probability questions "in reverse".  For example, we might ask "If I flip 100 coins, I can be 95% sure that I will get at least how many tails?".  We could answer this via guess-and-check with `pbinom()`, but fortunately, there is a function for it in **R** called `qbinom()`.


```{r qbinom}
textInput("n3",
          "How many trials do you have?",
          value = "0",
          width = '80%', placeholder = NULL)


textInput("p3",
          "What is the probability of success at each trial?",
          value = "0",
          width = '80%', placeholder = NULL)

textInput("q",
          "We want to find the number of successes that has a probability of ____ below it.",
          value = "0",
          width = '80%', placeholder = NULL)

bs_1_3 = reactive("qbinom(%ni, size = %ni, prob = %ni)")

```

```{r, context="server"}

output$code_3 <- renderText(
        makePrintText(
          base_string = bs_1_3(),
          old_input = c(),
          new_input = c(input$q, input$n3, input$p3)
        )
  )

output$output_3 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_1_3(), 
      old_input = c(),
      new_input = c(input$q, input$n3, input$p3)
    )
   ))
})

output$binom_var_3 <- renderText(
        makePrintText(
          base_string = "X ~ Binomial(%ni, %ni) \n P(X ≤ ?) = %ni",
          old_input = c(),
          new_input = c(input$n2, input$p2, input$k2)
        )
  )


output$distn_2 <- renderPlot(
  illustrate_binom(input$n2, input$p2, x = input$k2)
)

```

Random variable and probability statement:
```{r}
htmlOutput("binom_var_3")
```

Code:
```{r}
htmlOutput("code_3")
```

Result:
```{r}
verbatimTextOutput("output_3")
```


Illustration:
```{r}
plotOutput("distn_3")
```

### Your turn


```{r ex_3, exercise = TRUE}

```

```{r q_3}
textInput("a_3.1",
           "I flip 30 fair coins.  There is a 90% chance that I get less than how many heads?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.1")

textInput("a_3.2",
           "16% of all Starburst candies in a pack are cherry flavored.  I can be 85% sure that I will get at least how many cherry flavored Starbursts in a pack of 12?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.2")

textInput("a_3.3",
           "About 10% of all humans are left-handed.  I think there is a 95% chance that fewer than how many of my 35 students are left-handed?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.3")

```

```{r, context = "server"}

output$check_3.1 <- renderText(
  if(input$a_3.1 == ""){
     ""
  }else if(round(input$a_3.1,2) == round(qbinom(0.9, 30, .5), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_3.2 <- renderText(
  if(input$a_3.2 == ""){
     ""
  }else if(round(input$a_3.2,2) == round(qbinom(.15, 12, 0.16), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_3.3 <- renderText(
  if(input$a_3.3 == ""){
     ""
  }else if(round(input$a_3.3,2) == round(qbinom(.95, 35, 0.1), 2)){
      correct
  }else{
    incorrect
  }
)
```



## Challenge:  Practice predicting code

A good habit to get into when coding is to "translate" your code to English in your head, and try to predict what the output will be.  This is how to figure out if you are truly asking the computer to calculate what you hoped.  (Remember - computers will almost always output *something*!  It is your responsibility as the scientist to make sure you have asked the right question.)

For the following, try to guess what the output would be.  Do **not** run the code; instead, try to read it as a probability question and think through the answer.


*Give answers to 2 decimals*


```{r q_4}
textInput("a_4.1",
           "dbinom(1, size = 1, prob = 0.7)",
           value = "0",
           width = '80%', placeholder = NULL)

htmlOutput("check_4.1")

textInput("a_4.2",
           "pbinom(60, size = 60, prob = 0.13)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_4.2")

textInput("a_4.4",
           "qbinom(1, size = 60, prob = 0.13)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_4.4")


textInput("a_4.3",
           "pbinom(50, size = 100, prob = 0.5)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.3")

textInput("a_4.5",
           "qbinom(0.5, size = 100, prob = 0.5)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_4.5")
```

```{r, context = "server"}

output$check_4.1 <- renderText(
  if(input$a_4.1 == ""){
     ""
  }else if(as.numeric(input$a_4.1) == 0.7){
      correct
  }else{
    incorrect
  }
)

output$check_4.2 <- renderText(
  if(input$a_4.2 == ""){
     ""
  }else if(as.numeric(input$a_4.2) == 1){
      correct
  }else{
    incorrect
  }
)

output$check_4.3 <- renderText(
  if(input$a_4.3 == ""){
     ""
  }else if(as.numeric(input$a_4.3) == 0.5){
      correct
  }else{
    incorrect
  }
)

output$check_2.4 <- renderText(
  if(as.numeric(input$a_2.4) == ""){
     ""
  }else if(input$a_2.4 == 60){
      correct
  }else{
    incorrect
  }
)

output$check_4.5 <- renderText(
  if(input$a_4.5 == ""){
     ""
  }else if(as.numeric(input$a_4.5) == 50){
      correct
  }else{
    incorrect
  }
)


```