---
title: "Lab 6: Chi-Square"
author: "YOUR NAME HERE"
date: "October 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
### YOU CAN IGNORE EVERYTHING IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)

source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/packageLoaders2.R")

require(ggplot2)
require(dplyr)
require(tidyr)
require(reshape2)
require(fmsb)

data("HairEyeColor")

HairEyeColor <- HairEyeColor %>% tbl_df()

hair_eye <- array(unlist(apply(HairEyeColor, 1, function(x) rep(x[1:3], times = as.integer(x[4])))), dim = c(3,sum(HairEyeColor$n)))

rm(HairEyeColor)

hair_eye = data.frame(t(hair_eye))
names(hair_eye) = c("Hair.Color", "Eye.Color", "Gender")

set.seed(131)
hair_eye <- hair_eye %>% sample_n(nrow(hair_eye))

```


The dataset `hair_eye` contains hair color, eye color, and gender for several Statistics students.  

Use this dataset to discover something about hair color, eye color, and/or gender.  Your discussion should be thorough but brief - your knitted html should be only 1-2 pages.

Your discussion should include:
- A general overview and summary of the dataset.
- A statement of at least one clear research question.
- A plot that addresses your research question. (Hint: See the "Categorical Variables" exercise.)
- A Chi-Squared test that addresses your research question.  You should clearly state your hypotheses and your conclusion.
- A discussion of conditional proportions that address your research question.
- A clear overall conclusion.

For extra credit, also include:
- A discussion involving odds.  (Up to 5 points)
- A Relative Risk test and interpretation.  (Up to 10 points.)

```{r}
## A code chunk
```



