---
title: "Lab Exercise: Nonparametric Tests"
output: learnr::tutorial
tutorial:
  id: "Lab_Nonparametric"
  version: 2.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Lab_Exercise_NonParametric.html")){
  file.remove("Lab_Exercise_NonParametric.html")
}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)

knitr::opts_chunk$set(echo = FALSE)

tags$style(type="text/css",
         ".shiny-output-error { visibility: hidden; }",
         ".shiny-output-error:before { visibility: hidden; }"
)


source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/makeStrings.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')

titanic <- titanic %>% mutate(
   Family.Aboard = Siblings.Spouses.Aboard + Parents.Children.Aboard,
     Survived = factor(Survived,
                    levels = c(0,1),
                    labels = c("Died", "Lived")),
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third")),
     Has.Parents = case_when(
      Parents.Children.Aboard > 0 ~ "Yes_Parents",
      Parents.Children.Aboard == 0 ~ "No_Parents"
    ),
  Has.Parents = factor(Has.Parents)
)

richkids <- titanic %>% filter(Age < 18, Passenger.Class == "First")
### %ni = new input (red), %oi = old input (blue)
```


## Introduction

We are now going to narrow our focus in our study of the `titanic` dataset.  Consider the line of code below.  

```
richkids <- titanic %>% 
  filter(Age < 18, Passenger.Class == "First") %>%
  mutate(
    Has.Parents = case_when(
      Parents.Children.Aboard > 0 ~ "Yes_Parents",
      Parents.Children.Aboard == 0 ~ "No_Parents"
    )
  )
```

Here we have narrowed down our dataset to only the children travelling in First Class.  We have also made a new variable called `Has.Parents`, showing whether or not the child was travelling with his or her parents.  Let's take a look at the dataset.

```{r}
richkids %>% summary()
```





## Sign Tests

We will now learn to ask **R** to do nonparametric hypothesis tests for us.  We will start with a sign test.  Use the exercise below to answer the subsequent questions.

```{r, echo=FALSE}
textInput("var_1", 
          "We would like to make claims about which variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("val_1", 
          "What value of this variable are we interested in?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("num_1", 
          "How many samples are in this dataset?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)




bs_1 <- reactive("num_lived <- sum(%oi$%ni == '%ni')
print(num_lived) 
1-pbinom(num_lived, size = %ni, p = 0.5)")

```

```{r, context="server"}

output$code_1 <- renderText(
        makePrintText(
          base_string = bs_1(),
          old_input = c("richkids"),
          new_input = c(input$var_1, input$val_1, input$num_1)
        )
  )

output$output_1 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_1(),
          old_input = c("richkids"),
          new_input = c(input$var_1, input$val_1, input$num_1)
    )
   ))
})


output$hypos_1 <- renderText(
        makePrintText(
          base_string = "H0: First class children are just as likely to have %ni as not.\n  Ha: First class children are more likely to have %ni.",
          old_input = c(),
          new_input = c(input$val_1, input$val_1)
        )
  )

```

Hypotheses:
```{r}
htmlOutput("hypos_1")
```

Code:
```{r}
htmlOutput("code_1")
```

Result from code:
```{r}
verbatimTextOutput("output_1")
```

### Practice

Use this data to test whether First Class children were more likely to live or die.  Round answers to the nearest 2 decimals.

```{r ex_1, exercise = TRUE}

```

```{r q_1}

textInput("a_1.0",
           "How many first class children survived?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.0")
# 
# textInput("a_1.1",
#            "How many?",
#            value = "",
#            width = '80%', placeholder = NULL)
# 
# htmlOutput("check_1.1")


# 
# textInput("a_1.2",
#            "How many degrees of freedom were there?",
#            value = "",
#            width = '80%', placeholder = NULL)
# 
# htmlOutput("check_1.2")

textInput("a_1.3",
           "What was the p-value of this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.3")

radioButtons("a_1.4",
           "At the 0.05 level, we would",
           choices = c(
             "Reject the null" = "reject",
             "Fail to reject the null" = "fail")
)

htmlOutput("check_1.4")

radioButtons("a_1.5",
           "At the 0.05 level, we conclude that",
           choices = c(
             "First class children are more likely to die." = "no1",
             "First class children are more likely to survive." = "yes",
             "There is not enough evidence to make a conclusion." = "no3"
             )
)

htmlOutput("check_1.5")


```

```{r, context = "server"}

output$check_1.0 <- renderText(
  if(input$a_1.0 == ""){
     ""
  }else if(round(as.numeric(input$a_1.0),2) == round(sum(richkids$Survived == "Lived"), 2)){
      correct
  }else{
    incorrect
  }
)
# 
# output$check_1.1 <- renderText(
#   if(input$a_1.1 == ""){
#      ""
#   }else if(round(as.numeric(input$a_1.1), 2) == round(test()$statistic, 2)){
#       correct
#   }else{
#     incorrect
#   }
# )
#     
# output$check_1.2 <- renderText(
#   if(input$a_1.2 == ""){
#      ""
#   }else if(as.numeric(input$a_1.2) == 886){
#       correct
#   }else{
#     incorrect
#   }
# )

output$check_1.3 <- renderText(
  if(input$a_1.3 == ""){
     ""
  }else if(round(as.numeric(input$a_1.3), 2) == round(1-pbinom(11, 12, p = 0.5), 3)){
      correct
  }else{
    incorrect
  }
)

output$check_1.4 <- renderText(
  if(input$a_1.4 == ""){
     ""
  }else if(input$a_1.4 == "reject"){
      correct
  }else{
    incorrect
  }
)

output$check_1.5 <- renderText(
  if(input$a_1.5 == ""){
     ""
  }else if(input$a_1.5 == "yes"){
      correct
  }else{
    incorrect
  }
)
```


## Wilcox test

Now explore using `wilcox.test()` to more nuanced nonparametric tests.

```{r, echo=FALSE}
textInput("var_2", 
          "We would like to make claims about which QUANTITATIVE variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("groups_2", 
          "What CATEGORICAL, BINARY variable determines the two groups we wish to compare?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


radioButtons("alt_hypo_2",
             "Under the alternate hypothesis, there is a location shift in the direction of",
             choices = c("Group 1 is greater than Group 2." = "greater", 
                         "Group 1 is less than Group 2." = "less",
                         "either direction." = "two.sided")
)




bs_2 <- reactive("with(%oi, wilcox.test(%ni ~ %ni, alternative = '%ni'))")

```

```{r, context="server"}

output$code_2 <- renderText(
        makePrintText(
          base_string = bs_2(),
          old_input = c("richkids"),
          new_input = c(input$var_2, input$groups_2, input$alt_hypo_2)
        )
  )

output$output_2 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_2(),
          old_input = c("richkids"),
          new_input = c(input$var_2, input$groups_2, input$alt_hypo_2)
    )
   ))
})

alt2 <- reactive(switch(input$alt_hypo_2, "two.sided" = "not equal to", "greater" = "greater than", "less" = "less than"))

groups <- reactive(eval(parse(text = paste0("levels(titanic$", input$groups_2, ")"))))

output$hypos_2 <- renderText(
        makePrintText(
          base_string = "H0: There is no location shift in %ni.\n  Ha: There is a location shift in %ni, in the direction that %ni is %ni %ni.",
          old_input = c(),
          new_input = c(input$var_2, input$var_2, groups()[1], alt2(), groups()[2])
        )
  )

```

Hypotheses:
```{r}
htmlOutput("hypos_2")
```

Code:
```{r}
htmlOutput("code_2")
```

Result from code:
```{r}
verbatimTextOutput("output_2")
```

### Practice

Use this data to test whether First Class children with parents are younger than First Class children without parents.  Round answers to the nearest 2 decimals, except round the p-value to the nearest 3 decimals.

```{r ex_2, exercise = TRUE}

```

```{r q_2}
# 
# textInput("a_2.1",
#            "What sample mean age did we observe for men?",
#            value = "",
#            width = '80%', placeholder = NULL)
# 
# htmlOutput("check_2.1")
# 
# textInput("a_2.2",
#            "What sample mean age did we observe for women?",
#            value = "",
#            width = '80%', placeholder = NULL)
# 
# htmlOutput("check_2.2")


textInput("a_2.3",
           "What was the test statistic for this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.3")



textInput("a_2.4",
           "What was the p-value of this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.4")


radioButtons("a_2.5",
           "At the 0.05 level, we conclude that",
           choices = c(
             "There is evidence that older children were more likely to have parents aboard." = "no1",
              "There is evidence that younger children were more likely to have parents aboard." = "yes",
              "There is not enough evidence to make a conclusion." = "no2")
)

htmlOutput("check_2.5")

# radioButtons("a_2.6",
#            "Will a 95% confidence interval for the true difference in mean ages contain 0?",
#            choices = c(
#              "No way to tell" = "na",
#              "Yes" = "yes",
#               "No" = "no")
# )
# 
# htmlOutput("check_2.6")


```

```{r, context = "server"}

test2 <- reactive(with(richkids, wilcox.test(Age ~ Has.Parents, exact = FALSE, alternative = "greater")))
# 
# output$check_2.1 <- renderText(
#   if(input$a_2.1 == ""){
#      ""
#   }else if(round(as.numeric(input$a_2.1),2) == round(mean(titanic$Age[titanic$Sex == "M"]), 2)){
#       correct
#   }else{
#     incorrect
#   }
# )
# 
# 
# output$check_2.2 <- renderText(
#   if(input$a_2.2 == ""){
#      ""
#   }else if(round(as.numeric(input$a_2.2),2) == round(mean(titanic$Age[titanic$Sex == "F"]), 2)){
#       correct
#   }else{
#     incorrect
#   }
# )

output$check_2.3 <- renderText(
  if(input$a_2.3 == ""){
     ""
  }else if(round(as.numeric(input$a_2.3), 2) == round(test2()$statistic, 2)){
      correct
  }else{
    incorrect
  }
)


output$check_2.4 <- renderText(
  if(input$a_2.4 == ""){
     ""
  }else if(round(as.numeric(input$a_2.4), 3) == round(test2()$p.value, 3)){
      correct
  }else{
    incorrect
  }
)

output$check_2.5 <- renderText(
  if(input$a_2.5 == ""){
     ""
  }else if(input$a_2.5 == "yes"){
      correct
  }else{
    incorrect
  }
)

# output$check_2.6 <- renderText(
#   if(input$a_2.6 == ""){
#      ""
#   }else if(input$a_2.6 == "no"){
#       correct
#   }else{
#     incorrect
#   }
# )
```


## Paired nonparametric tests

Finally, we will also use `wilcox.test()` to perform nonparametric tests of paired data.

```{r, echo=FALSE}
textInput("var_3_1", 
          "Which QUANTITATIVE variable is our first measurement?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("var_3_2", 
          "Which QUANTITATIVE variable is our second measurement?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


radioButtons("alt_hypo_3",
             "Under the alternate hypothesis, there is a location shift in the direction of",
             choices = c("Variable 1 is greater than Variable 2." = "greater", 
                         "Variable 1 is less than Variable 2." = "less",
                         "either direction." = "two.sided")
)




bs_3 <- reactive("with(%oi, wilcox.test(%ni, %ni, alternative = '%ni', %oi))")

```

```{r, context="server"}

output$code_3 <- renderText(
        makePrintText(
          base_string = bs_3(),
          old_input = c("richkids", "paired = TRUE"),
          new_input = c(input$var_3_1, input$var_3_2, input$alt_hypo_3)
        )
  )

output$output_3 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_3(),
          old_input = c("richkids", "paired = TRUE"),
          new_input = c(input$var_3_1, input$var_3_2, input$alt_hypo_3)
    )
   ))
})

alt3 <- reactive(switch(input$alt_hypo_3, "two.sided" = "not equal to", "greater" = "greater than", "less" = "less than"))

output$hypos_3 <- renderText(
        makePrintText(
          base_string = "H0: There is no location shift between %ni and %ni.\n  Ha: There is a location shift, in the direction that %ni is %ni %ni.",
          old_input = c(),
          new_input = c(input$var_3_1, input$var_3_2, input$var_3_1, alt3(), input$var_3_2)
        )
  )

```

Hypotheses:
```{r}
htmlOutput("hypos_3")
```

Code:
```{r}
htmlOutput("code_3")
```

Result from code:
```{r}
verbatimTextOutput("output_3")
```

### Practice

Use this data to test whether the First Class children on the Titanic had more Siblings/Spouses aboard or more Parents/Children aboard.  Round answers to the nearest 2 decimals, except round the p-value to the nearest 3 decimals.

```{r ex_3, exercise = TRUE}

```

```{r q_3}
# 
# textInput("a_3.1",
#            "What sample mean did we observe for the difference between Siblings/Spouses aboard and Parents/Children aboard?",
#            value = "",
#            width = '80%', placeholder = NULL)
# 
# htmlOutput("check_3.1")

# textInput("a_3.2",
#            "What sample mean did we observe for the number of Parents/Children?",
#            value = "",
#            width = '80%', placeholder = NULL)
# 
# htmlOutput("check_3.2")


textInput("a_3.3",
           "What was the test statistic for this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.3")



textInput("a_3.4",
           "What was the p-value of this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.4")


radioButtons("a_3.5",
           "At the 0.05 level, we conclude that",
           choices = c(
               "There is not enough evidence to make a conclusion." = "no3",
             "There is evidence that First Class children had more Siblings/Spouses aboard." = "no2",
              "There is evidence that First Class children had more Parents/Children aboard." = "yes"
           )
)

htmlOutput("check_3.5")

# radioButtons("a_3.6",
#            "Will a 95% confidence interval for the true mean difference of Siblings/Spouses versus Parents/Children contain 0?",
#            choices = c(
#              "No way to tell" = "na",
#              "Yes" = "yes",
#               "No" = "no")
# )
# 
# htmlOutput("check_3.6")


```

```{r, context = "server"}

test3 <- reactive(with(richkids, wilcox.test(Siblings.Spouses.Aboard, Parents.Children.Aboard, alternative = "two.sided", paired = TRUE, exact = FALSE)))

# output$check_3.1 <- renderText(
#   if(input$a_3.1 == ""){
#      ""
#   }else if(round(as.numeric(input$a_3.1),2) == round(mean(titanic$Siblings.Spouses.Aboard - titanic$Parents.Children.Aboard), 2)){
#       correct
#   }else{
#     incorrect
#   }
# )


# output$check_3.2 <- renderText(
#   if(input$a_3.2 == ""){
#      ""
#   }else if(round(as.numeric(input$a_3.2),2) == round(mean(titanic$Parents.Children.Aboard), 2)){
#       correct
#   }else{
#     incorrect
#   }
# )

output$check_3.3 <- renderText(
  if(input$a_3.3 == ""){
     ""
  }else if(round(as.numeric(input$a_3.3), 2) == round(test3()$statistic, 2)){
      correct
  }else{
    incorrect
  }
)


output$check_3.4 <- renderText(
  if(input$a_3.4 == ""){
     ""
  }else if(round(as.numeric(input$a_3.4), 3) == round(test3()$p.value, 3)){
      correct
  }else{
    incorrect
  }
)

output$check_3.5 <- renderText(
  if(input$a_3.5 == ""){
     ""
  }else if(input$a_3.5 == "yes"){
      correct
  }else{
    incorrect
  }
)

# output$check_3.6 <- renderText(
#   if(input$a_3.6 == ""){
#      ""
#   }else if(input$a_3.6 == "no"){
#       correct
#   }else{
#     incorrect
#   }
# )
```
