---
title: "Lab 1: Looking at Datasets"
author: "YOUR NAME HERE"
date: "September 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
### YOU CAN IGNORE EVERYTHING IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE, fig.width=5, fig.height=5)

source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/packageLoaders2.R")

require(ggplot2)
require(dplyr)
require(tidyr)

edu = read.csv("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/Edu-Data.csv")

```

## Instructions

This analysis is an exploration of a dataset containing information about grade school student educational performance in the Middle East.  The data has been loaded for you above, and named `edu`.

In the following code chunk, look at the first few rows of `edu`.  You can also reference the file "Edu_data_description.txt" that is linked on PolyLearn, to read about this dataset.

```{r}

```

Our goal will be to determine which aspects of a student's experience most relate to his or her performance.  The variable `Class` gives information about student performance - either "High", "Medium", or "Low".

Choose two other **categorical** variables that you would like to study.  Use the code chunk below to `select` from your dataset only Class and your two chosen variables.

```{r}

```


## The Education Dataset


[In this section, use vocabulary from the course to describe your new dataset.]

## Research Questions

[In this section, give 2-3 research questions about student class performance that you will address with this data.]

## Plotting single variables

Use the code chunk below to create an appropriate plot for each of your three categorical variables.

```{r}

```

[In this section, discuss your first plot.]

```{r}

```

[In this section, discuss your second plot.]

```{r}

```

[In this section, discuss your third plot.]

## Answering the research questions

Using code chunks to make plots, combined with discussions of each plot, address your research questions with data.


