---
title: "Lab Exercise: t-tests and Confidence Intervals"
output: learnr::tutorial
tutorial:
  id: "Lab_ttests"
  version: 1.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Lab_Exercise_t_tests.html")){
  file.remove("Lab_Exercise_t_tests.html")
}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)

knitr::opts_chunk$set(echo = FALSE)

tags$style(type="text/css",
         ".shiny-output-error { visibility: hidden; }",
         ".shiny-output-error:before { visibility: hidden; }"
)


source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/makeStrings.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')

titanic <- titanic %>% mutate(
   Family.Aboard = Siblings.Spouses.Aboard + Parents.Children.Aboard,
     Survived = factor(Survived,
                    levels = c(0,1),
                    labels = c("Died", "Lived")),
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)
### %ni = new input (red), %oi = old input (blue)
```


## Introduction

We continue our journey through the `titanic` dataset.  The data has already been loaded for you, and cleaned.  Remind yourself what the dataset looks like, and what variables are present

```{r, echo = TRUE}
titanic %>% head()
titanic %>% summary()
```


## One sample t-tests

We will now learn to ask **R** to do the hard work in doing hypothesis tests for us.  The function `t.test()` will let us compute t-scores and p-values automatically.  Try it below.

```{r, echo=FALSE}
textInput("var_1", 
          "We would like to make claims about the TRUE mean of which QUANTITATIVE variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("null_1", 
          "Under the null, what is the true mean equal to?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


radioButtons("alt_hypo_1",
             "Under the alternate hypothesis, the true mean is",
             choices = c("greater than the null mean." = "greater", 
                         "less than the null mean." = "less",
                         "not equal to the null mean." = "two.sided")
)




bs_1 <- reactive("with(%oi, t.test(%ni, mu = %ni, alternative = '%ni'))")

```

```{r, context="server"}

output$code_1 <- renderText(
        makePrintText(
          base_string = bs_1(),
          old_input = c("titanic"),
          new_input = c(input$var_1, input$null_1, input$alt_hypo_1)
        )
  )

output$output_1 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_1(),
          old_input = c("titanic"),
          new_input = c(input$var_1, input$null_1, input$alt_hypo_1)
    )
   ))
})

alt1 <- reactive(switch(input$alt_hypo_1, "two.sided" = "not equal to", "greater" = "greater than", "less" = "less than"))

#df <- reactive(eval(parse(paste0("length(titanic$", input$var_1, ") - 1"))))

output$hypos_1 <- renderText(
        makePrintText(
          base_string = "H0: The true mean of %ni is equal to %ni.\n  Ha: The true mean of %ni is %ni %ni.",
          old_input = c(),
          new_input = c(input$var_1, input$null_1, input$var_1, alt1(), input$null_1)
        )
  )

```

Hypotheses:
```{r}
htmlOutput("hypos_1")
```

Code:
```{r}
htmlOutput("code_1")
```

Result from code:
```{r}
verbatimTextOutput("output_1")
```

### Practice

Use this data to test whether the mean age on the Titanic was below 30.  Round answers to the nearest 2 decimals.

```{r ex_1, exercise = TRUE}

```

```{r q_1}

textInput("a_1.0",
           "What sample mean did we observe?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.0")

textInput("a_1.1",
           "What was the t-score for this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1")



textInput("a_1.2",
           "How many degrees of freedom were there?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.2")

textInput("a_1.3",
           "What was the p-value of this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.3")

radioButtons("a_1.4",
           "At the 0.05 level, we would",
           choices = c(
             "Reject the null" = "reject",
             "Fail to reject the null" = "fail")
)

htmlOutput("check_1.4")

radioButtons("a_1.5",
           "At the 0.05 level, we conclude that",
           choices = c(
             "The true mean age of all Titanic passengers is exactly 29.47." = "no1",
             "The true mean age of all Titanic passengers is exactly 30." = "no2",
             "There is not strong evidence that the true mean age of all Titanic passengers is exactly 30." = "no3",
             "There is not strong evidence that the true mean age of all Titanic passengers is less than 30." = "yes",
             "There is not strong evidence that the true mean age of all Titanic passengers is not equal to 30." = "no4")
)

htmlOutput("check_1.5")


```

```{r, context = "server"}

test <- reactive(t.test(titanic$Age, mu = 30, alternative = "less"))

output$check_1.0 <- renderText(
  if(input$a_1.0 == ""){
     ""
  }else if(round(as.numeric(input$a_1.0),2) == round(mean(titanic$Age), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.1 <- renderText(
  if(input$a_1.1 == ""){
     ""
  }else if(round(as.numeric(input$a_1.1), 2) == round(test()$statistic, 2)){
      correct
  }else{
    incorrect
  }
)
    
output$check_1.2 <- renderText(
  if(input$a_1.2 == ""){
     ""
  }else if(as.numeric(input$a_1.2) == 886){
      correct
  }else{
    incorrect
  }
)

output$check_1.3 <- renderText(
  if(input$a_1.3 == ""){
     ""
  }else if(round(as.numeric(input$a_1.3), 2) == round(test()$p.value, 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.4 <- renderText(
  if(input$a_1.4 == ""){
     ""
  }else if(input$a_1.4 == "fail"){
      correct
  }else{
    incorrect
  }
)

output$check_1.5 <- renderText(
  if(input$a_1.5 == ""){
     ""
  }else if(input$a_1.5 == "yes"){
      correct
  }else{
    incorrect
  }
)
```


## Two-sample t-tests

Now explore using `t.test()` to perform two-sample hypothesis tests.

```{r, echo=FALSE}
textInput("var_2", 
          "We would like to make claims about the TRUE means of which QUANTITATIVE variable?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("groups_2", 
          "What CATEGORICAL, BINARY variable determines the two groups we wish to compare means of?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


radioButtons("alt_hypo_2",
             "Under the alternate hypothesis, the true mean of Group 1 is",
             choices = c("greater than the true mean of Group 2." = "greater", 
                         "less than the true mean of Group 2." = "less",
                         "not equal to the true mean of Group 2." = "two.sided")
)




bs_2 <- reactive("with(%oi, t.test(%ni ~ %ni, alternative = '%ni'))")

```

```{r, context="server"}

output$code_2 <- renderText(
        makePrintText(
          base_string = bs_2(),
          old_input = c("titanic"),
          new_input = c(input$var_2, input$groups_2, input$alt_hypo_2)
        )
  )

output$output_2 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_2(),
          old_input = c("titanic"),
          new_input = c(input$var_2, input$groups_2, input$alt_hypo_2)
    )
   ))
})

alt2 <- reactive(switch(input$alt_hypo_2, "two.sided" = "not equal to", "greater" = "greater than", "less" = "less than"))

groups <- reactive(eval(parse(text = paste0("levels(titanic$", input$groups_2, ")"))))

output$hypos_2 <- renderText(
        makePrintText(
          base_string = "H0: The true mean %ni of %nis is equal to the true mean %ni of %nis.\n  Ha: The true mean %ni of %nis is %ni the true mean %ni of %nis.",
          old_input = c(),
          new_input = c(input$var_2, groups()[1], input$var_2, groups()[2], input$var_2, groups()[1], alt2(), input$var_2, groups()[2])
        )
  )

```

Hypotheses:
```{r}
htmlOutput("hypos_2")
```

Code:
```{r}
htmlOutput("code_2")
```

Result from code:
```{r}
verbatimTextOutput("output_2")
```

### Practice

Use this data to test whether the mean age on the Titanic was the same for men and women.  Round answers to the nearest 2 decimals, except round the p-value to the nearest 3 decimals.

```{r ex_2, exercise = TRUE}

```

```{r q_2}

textInput("a_2.1",
           "What sample mean age did we observe for men?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.1")

textInput("a_2.2",
           "What sample mean age did we observe for women?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.2")


textInput("a_2.3",
           "What was the t-score for this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.3")



textInput("a_2.4",
           "What was the p-value of this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_2.4")


radioButtons("a_2.5",
           "At the 0.05 level, we conclude that",
           choices = c(
             "There is evidence that men and women on the Titanic had the same mean age." = "no1",
              "There is not enough evidence that men and women on the Titanic had the same mean age." = "no2",
              "There is evidence that men and women on the Titanic had different mean ages." = "yes",
             "There is not enough evidence that men and women on the Titanic had different mean ages." = "no3")
)

htmlOutput("check_2.5")

radioButtons("a_2.6",
           "Will a 95% confidence interval for the true difference in mean ages contain 0?",
           choices = c(
             "No way to tell" = "na",
             "Yes" = "yes",
              "No" = "no")
)

htmlOutput("check_2.6")


```

```{r, context = "server"}

test2 <- reactive(t.test(Age ~ Sex, data = titanic, alternative = "two.sided"))

output$check_2.1 <- renderText(
  if(input$a_2.1 == ""){
     ""
  }else if(round(as.numeric(input$a_2.1),2) == round(mean(titanic$Age[titanic$Sex == "M"]), 2)){
      correct
  }else{
    incorrect
  }
)


output$check_2.2 <- renderText(
  if(input$a_2.2 == ""){
     ""
  }else if(round(as.numeric(input$a_2.2),2) == round(mean(titanic$Age[titanic$Sex == "F"]), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_2.3 <- renderText(
  if(input$a_2.3 == ""){
     ""
  }else if(round(as.numeric(input$a_2.3), 2) == round(test2()$statistic, 2)){
      correct
  }else{
    incorrect
  }
)


output$check_2.4 <- renderText(
  if(input$a_2.4 == ""){
     ""
  }else if(round(as.numeric(input$a_2.4), 3) == round(test2()$p.value, 3)){
      correct
  }else{
    incorrect
  }
)

output$check_2.5 <- renderText(
  if(input$a_2.5 == ""){
     ""
  }else if(input$a_2.5 == "yes"){
      correct
  }else{
    incorrect
  }
)

output$check_2.6 <- renderText(
  if(input$a_2.6 == ""){
     ""
  }else if(input$a_2.6 == "no"){
      correct
  }else{
    incorrect
  }
)
```


## Paired t-tests

Now explore using `t.test()` to perform hypothesis tests of paired data.

```{r, echo=FALSE}
textInput("var_3_1", 
          "Which QUANTITATIVE variable is our first measurement?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)

textInput("var_3_2", 
          "Which QUANTITATIVE variable is our second measurement?", 
          value = "", 
          width = '80%', 
          placeholder = NULL)


radioButtons("alt_hypo_3",
             "Under the alternate hypothesis, the true mean difference is",
             choices = c("greater than 0." = "greater", 
                         "less than 0." = "less",
                         "not equal to 0." = "two.sided")
)




bs_3 <- reactive("with(%oi, t.test(%ni, %ni, alternative = '%ni', %oi))")

```

```{r, context="server"}

output$code_3 <- renderText(
        makePrintText(
          base_string = bs_3(),
          old_input = c("titanic", "paired = TRUE"),
          new_input = c(input$var_3_1, input$var_3_2, input$alt_hypo_3)
        )
  )

output$output_3 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_3(),
          old_input = c("titanic", "paired = TRUE"),
          new_input = c(input$var_3_1, input$var_3_2, input$alt_hypo_3)
    )
   ))
})

alt3 <- reactive(switch(input$alt_hypo_3, "two.sided" = "not equal to", "greater" = "greater than", "less" = "less than"))

output$hypos_3 <- renderText(
        makePrintText(
          base_string = "H0: The true mean difference betweeen %ni and %ni is equal to 0.\n  Ha: The true mean difference of %ni and %ni is %ni 0.",
          old_input = c(),
          new_input = c(input$var_3_1, input$var_3_2, input$var_3_1, input$var_3_2, alt3())
        )
  )

```

Hypotheses:
```{r}
htmlOutput("hypos_3")
```

Code:
```{r}
htmlOutput("code_3")
```

Result from code:
```{r}
verbatimTextOutput("output_3")
```

### Practice

Use this data to test whether the passengers on the Titanic had more Siblings/Spouses aboard or more Parents/Children aboard.  Round answers to the nearest 2 decimals, except round the p-value to the nearest 3 decimals.

```{r ex_3, exercise = TRUE}

```

```{r q_3}

textInput("a_3.1",
           "What sample mean did we observe for the difference between Siblings/Spouses aboard and Parents/Children aboard?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.1")

# textInput("a_3.2",
#            "What sample mean did we observe for the number of Parents/Children?",
#            value = "",
#            width = '80%', placeholder = NULL)
# 
# htmlOutput("check_3.2")


textInput("a_3.3",
           "What was the t-score for this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.3")



textInput("a_3.4",
           "What was the p-value of this test?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.4")


radioButtons("a_3.5",
           "At the 0.05 level, we conclude that",
           choices = c(
               "There is not enough evidence to make a conclusion." = "no3",
             "There is evidence that passengers had more Siblings/Spouses aboard." = "yes",
              "There is evidence that passengers had more Parents/Children aboard." = "no2"
           )
)

htmlOutput("check_3.5")

radioButtons("a_3.6",
           "Will a 95% confidence interval for the true mean difference of Siblings/Spouses versus Parents/Children contain 0?",
           choices = c(
             "No way to tell" = "na",
             "Yes" = "yes",
              "No" = "no")
)

htmlOutput("check_3.6")


```

```{r, context = "server"}

test3 <- reactive(with(titanic, t.test(Siblings.Spouses.Aboard, Parents.Children.Aboard, alternative = "two.sided", paired = TRUE)))

output$check_3.1 <- renderText(
  if(input$a_3.1 == ""){
     ""
  }else if(round(as.numeric(input$a_3.1),2) == round(mean(titanic$Siblings.Spouses.Aboard - titanic$Parents.Children.Aboard), 2)){
      correct
  }else{
    incorrect
  }
)


# output$check_3.2 <- renderText(
#   if(input$a_3.2 == ""){
#      ""
#   }else if(round(as.numeric(input$a_3.2),2) == round(mean(titanic$Parents.Children.Aboard), 2)){
#       correct
#   }else{
#     incorrect
#   }
# )

output$check_3.3 <- renderText(
  if(input$a_3.3 == ""){
     ""
  }else if(round(as.numeric(input$a_3.3), 2) == round(test3()$statistic, 2)){
      correct
  }else{
    incorrect
  }
)


output$check_3.4 <- renderText(
  if(input$a_3.4 == ""){
     ""
  }else if(round(as.numeric(input$a_3.4), 3) == round(test3()$p.value, 3)){
      correct
  }else{
    incorrect
  }
)

output$check_3.5 <- renderText(
  if(input$a_3.5 == ""){
     ""
  }else if(input$a_3.5 == "yes"){
      correct
  }else{
    incorrect
  }
)

output$check_3.6 <- renderText(
  if(input$a_3.6 == ""){
     ""
  }else if(input$a_3.6 == "no"){
      correct
  }else{
    incorrect
  }
)
```
