---
title: "Lab Exercise: The t-Distribution"
output: learnr::tutorial
tutorial:
  id: "Lab_tDist"
  version: 1.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Lab_Exercise_tDist.html")){
  file.remove("Lab_Exercise_tDist.html")
}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)

knitr::opts_chunk$set(echo = FALSE)


source("https://raw.githubusercontent.com/kbodwin/Introductory-Statistics-Labs/master/Scripts/makeStrings.R")
source("https://raw.githubusercontent.com/kbodwin/Introductory-Statistics-Labs/master/Scripts/plotDensities.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

```


## Introduction

In this exercise, we will be using **R** to calculate information about *t-scores* and the *t-distribution*.  

## Area below a value

Recall that for Normal random variables, `pnorm()` calculated the probability of getting less than or equal to a certain number of successes.  Similarly, `pt()` will find the probabilities for a t-distribution.  There are two extra complications, however:

1. For a t-distribution, you must also supply the **degrees of freedom**.
2. `pt()` only calculates areas on the **standard curve**; that is, areas below **t-scores**.  Thus, you must first convert the value you are interested in to a t-score.

Try it below.


```{r tscore}
textInput("mu1",
          "What is the mean of your random variable?",
          value = "0",
          width = '80%', placeholder = NULL)


textInput("sigma1",
          "What is the standard deviation of your random variable?",
          value = "1",
          width = '80%', placeholder = NULL)

textInput("x1",
          "What cutoff value do you want to find the area below?",
          value = "0",
          width = '80%', placeholder = NULL)

bs_0 = reactive("t_score <- (%ni - %ni)/%ni \n t_score")


```


```{r, context="server"}

output$code_0 <- renderText(
        makePrintText(
          base_string = bs_0(),
          old_input = c(),
          new_input = c(input$x1, input$mu1, input$sigma1)
        )
  )

output$output_0 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_0(),
          old_input = c(),
          new_input = c(input$x1, input$mu1, input$sigma1)
    )
   ))
})


output$t_score_0 <- renderText(
        makePrintText(
          base_string = "What is t-score of %ni, if X has a mean of %ni and a standard deviation of %ni?",
          old_input = c(),
          new_input = c(input$x1, input$mu1, input$sigma1)
        )
  )


```

Summary sentence:
```{r}
htmlOutput("t_score_0")
```

Code:
```{r}
htmlOutput("code_0")
```

Result from code:
```{r}
verbatimTextOutput("output_0")
```



```{r pt}
textInput("t1",
          "What is the t-score you calculated above?",
          value = "",
          width = '80%', placeholder = NULL)

textInput("df1",
          "How many degrees of freedom do you have?",
          value = "",
          width = '80%', placeholder = NULL)

bs_1 = reactive("pt(%ni, df = %ni)")

```

```{r, context="server"}

output$code_1 <- renderText(
        makePrintText(
          base_string = bs_1(),
          old_input = c(),
          new_input = c(input$t1, input$df1)
        )
  )

output$output_1 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_1(),
          old_input = c(),
          new_input = c(input$t1, input$df1)
    )
   ))
})

output$t_var_1 <- renderText(
        makePrintText(
          base_string = "The t-score is t-distributed with df = %ni \n P(t ≤ %ni) = ?",
          old_input = c(),
          new_input = c(input$df1, input$t1)
        )
  )


output$distn_1 <- renderPlot(
  illustrate_t(df = input$df1, t_score = input$t1)
)

```

Random variable and probability statement:
```{r}
htmlOutput("t_var_1")
```

Code:
```{r}
htmlOutput("code_1")
```

Result from code:
```{r}
verbatimTextOutput("output_1")
```


Illustration:
```{r}
plotOutput("distn_1")
```

## Practice with `pt()`

In the box below, you may run any line of **R** code.  Use the box, and/or the exercise above, to answer the following questions.


```{r ex_1.1, exercise = TRUE}

```


*Round your answers to 2 decimals*

```{r q_1}

textInput("a_1.1",
           "What is the area below a t-score of -1.38 with 25 degrees of freedom?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1")

```

Suppose you sample 18 people and record their heights.  You find a sample mean of 67 inches and a sample standard deviation of 5 inches.  Use the code box below to calculate your answers to the following questions.  (DRAW PICTURES!)

```{r q_1.2}


textInput("a_1.2",
           "What would be your p-value for testing the alternate hypothesis \"The true average height of humans is more than 65 inches\"?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.2")

textInput("a_1.3",
           "If you had used a Normal distribution, rather than the t-distribution, what would your p-value have been?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.3")


```

```{r, context = "server"}

output$check_1.2 <- renderText(
  if(input$a_1.2 == ""){
     ""
  }else if(round(as.numeric(input$a_1.2),2) == round(1-pt(2*sqrt(18)/5, df = 17), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.3 <- renderText(
  if(input$a_1.3 == ""){
     ""
  }else if(round(as.numeric(input$a_1.3), 2) == round(1-pnorm(67, mean = 65, sd = 5/sqrt(18)), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.1 <- renderText(
  if(input$a_1.1 == ""){
     ""
  }else if(round(as.numeric(input$a_1.1), 2) == round(pt(-1.38, df = 18), 2)){
      correct
  }else{
    incorrect
  }
)
```




## Finding t-star values

Recall that we used `qnorm()` to find cutoff values with a given area under the curve.  Similarly, `qt()` will give us the t-score with a certain area below it, for a given degrees of freedom.  This is useful for creating confidence intervals, because it is how we find the $t^*$ value for a particular confidence level and degrees of freedom.


```{r qnorm}
textInput("df2",
          "How many degrees of freedom do you have?",
          value = "0",
          width = '80%', placeholder = NULL)


textInput("q2",
          "We want to find the t-score that has an area of ____ below it.",
          value = "0",
          width = '80%', placeholder = NULL)

bs_3 = reactive("qt(%ni, df = %ni)")

```

```{r, context="server"}

output$code_3 <- renderText(
        makePrintText(
          base_string = bs_3(),
          old_input = c(),
          new_input = c(input$q2, input$df2)
        )
  )

output$output_3 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_3(), 
      old_input = c(),
     new_input = c(input$q2, input$df2)
    )
   ))
})

output$t_var_3 <- renderText(
        makePrintText(
          base_string = "What t-score with %ni degrees of freedom has an area of %ni below it?",
          old_input = c(),
          new_input = c(input$q2, input$df2)
        )
  )


output$distn_3 <- renderPlot(
  illustrate_t(df = input$df2, q = input$q2)
)

```

Random variable question:
```{r}
htmlOutput("t_var_3")
```

Code:
```{r}
htmlOutput("code_3")
```

Result:
```{r}
verbatimTextOutput("output_3")
```


Illustration:
```{r}
plotOutput("distn_3")
```


## Practice with `qt()`

Use code to answer the following questions.

```{r ex_3, exercise = TRUE}

```

```{r q_3}
textInput("a_3.1",
           "What would be our t* value for a 95% confidence interval with 90 degrees of freedom?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.1")

textInput("a_3.2",
           "What would be our t* value for a 95% confidence interval with 27 degrees of freedom?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.2")

textInput("a_3.3",
           "What would be our t* value for a 90% confidence interval with 27 degrees of freedom?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.3")

textInput("a_3.4",
           "You decide to perform a one-sided hypothesis test at the 0.05 level based on a sample mean from 28 samples.  What t-score is the cutoff for rejecting your null hypothesis?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.4")

textInput("a_3.5",
           "You decide to perform a two-sided hypothesis test at the 0.05 level based on a sample mean from 28 samples.  What (positive) t-score is the cutoff for rejecting your null hypothesis?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.5")

```

```{r, context = "server"}

output$check_3.1 <- renderText(
  if(input$a_3.1 == ""){
     ""
  }else if(round(as.numeric(input$a_3.1),2) == round(qt(.975, df = 90), 2)){
      correct
  }else if(round(as.numeric(input$a_3.2),2) == round(qt(.95, df = 90), 2)){
    "<font color='red'>Not quite... what area do we want below the t* value?</font><br><br>"
  }else{
    incorrect
  }
)

output$check_3.2 <- renderText(
  if(input$a_3.2 == ""){
     ""
  }else if(round(as.numeric(input$a_3.2),2) == round(qt(.975, df = 27), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_3.3 <- renderText(
  if(input$a_3.3 == ""){
     ""
  }else if(round(as.numeric(input$a_3.3),2) == round(qt(.95, df = 27), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_3.4 <- renderText(
  if(input$a_3.4 == ""){
     ""
  }else if(round(as.numeric(input$a_3.4),2) == round(qt(.95, df = 27), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_3.4 <- renderText(
  if(input$a_3.4 == ""){
     ""
  }else if(round(as.numeric(input$a_3.4),2) == round(qt(.975, df = 27), 2)){
      correct
  }else{
    incorrect
  }
)
```
