---
title: "Lab 5: "
author: "YOUR NAME HERE"
date: "October 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
### YOU CAN IGNORE EVERYTHING IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)

source("https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Scripts/packageLoaders2.R")

require(ggplot2)
require(dplyr)
require(tidyr)
require(reshape2)


data(french_fries)
french_fries <- french_fries %>% filter(subject == 10, rep == 1, treatment != "3") %>% select(-subject, -rep, -time)

```


## Description of Dataset

Today's dataset is about French Fries.  Mmmmm.  Here is what the authors of the dataset have to say about it:

```
This data was collected from a sensory experiment conducted at Iowa State University in 2004. The investigators were interested in the effect of using two different fryer oils had on the taste of the fries.

Variables:

treatment (type of oil),

potato-y flavour,

buttery flavour,

grassy flavour,

rancid flavour,

painty flavour
```

```{r}
## Explore the data
```

[Briefly discuss this dataset here]

## Potatoey-ness

Suppose we wish to study the differences between the two oil treatments.

### 1

Conduct an appropriate test for whether Oil Treatment 1 results in fries that are more or less "potatoey" than Oil Treatment 2.  You may assume, for now, that "potato-ness" is Normally distributed.

```{r}
## Your test here
```

[Discuss your test here.]

### 2

Run the following code to make a QQ plot for the "potato-ness" of a fry.  Then add to it to also make a histogram of potatoeyness.

```{r}
ggplot(french_fries, aes(sample = potato)) + geom_qq()

# your code for a histogram
```
Comment on this QQ plot and histogram.  What did you learn?


### 3

Now assume that "potato-ness" is *not* Normally distributed.  Conduct an appropriate test for whether Oil Treatment 1 results in fries that are more or less "potatoey" than Oil Treatment 2. 


```{r}
## Your test here
```

[Discuss your test here.]

## Do potatoes taste like potatoes?

It seems very strange to me that a french fry, which is made of a potato, would not be rated as tasting like "potato".  Let's look at whether or not each fry was rated as potatoey.

```{r}
french_fries <- french_fries %>% mutate(
  is.potato = potato > 0
)
```

### 1

Alter the code below to use `count()` to find out how many fries were rated as potatoey.

```{r}
french_fries %>% count()
```

There were [-----] fries rated "potato" and [----] fries not rated "potato"

### 2

If fries are equally likely to be rated potatoey as not, what is the distribution of the random variable

X = number of fries out of 20 that are rated as potatoey


[Your answer here]

### 3

If fries are equally likely to be rated potatoey as not, how unlikely is it to see as many fries (or more) rated "potato" as we saw in this dataset?

```{r}
# your calculation here
```

[Discuss your answer.  What did you learn about potatoeyness of fries?]


## Grassy or Painty?

Is a fry from Oil Treatment 1 more likely to be rated higher on the "painty" score or the "grassy" score?  Let's filter our dataset down to only Treatment 1 to find out.


```{r}
fries_1 <- french_fries %>% filter(treatment == 1)

```

### 1

Assume ratings of grassyness and paintyness are Normally distributed, and do an appropriate test.

```{r}

# your test here
```

[Discuss your test here.]

### 2

Make a qqplot and histogram, and comment on them.

```{r}
# Your code for plots here
```


### 3
Now do NOT assume Normality, and do an appropriate test.

```{r}

# your test here
```

[Discuss your test here.]


## Rancid Fries???

Discuss the relationship between how "rancid" a fry is rated, and which oil treatment was used.  Be sure to justify your approach, interpret your results, and create at least one plot to support your answer.

```{r}
# Your code here
```

[Your discussion here.]

