---
title: "Lab Exercise: Normal Random Variables"
output: learnr::tutorial
tutorial:
  id: "Lab_Normal"
  version: 1.0
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

if(file.exists("Lab_Exercise_Normal.html")){
  file.remove("Lab_Exercise_Normal.html")
}

require(learnr)
require(ggplot2)
require(dplyr)
require(shiny)
require(tidyr)
require(httpuv)

knitr::opts_chunk$set(echo = FALSE)


source("https://raw.githubusercontent.com/kbodwin/Introductory-Statistics-Labs/master/Scripts/makeStrings.R")
source("https://raw.githubusercontent.com/kbodwin/Introductory-Statistics-Labs/master/Scripts/plotDensities.R")

correct <- "<font color='red'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Not quite right...</font><br><br>"
congrats <- "<font color='red'>You did it!</font><br><br>"

```


## Introduction

In this exercise, we will be using **R** to calculate information about Normal random variables.  

## Area below a value

Recall that for Binomial random variables, `pbinom()` calculated the probability of getting less than or equal to a certain number of successes.  Similarly, `pnorm()` will find the probability of a Normal random variable taking a value less than or equal to a given cutoff.


```{r pnorm}
textInput("mu1",
          "What is the mean of your Normal random variable?",
          value = "0",
          width = '80%', placeholder = NULL)


textInput("sigma1",
          "What is the standard deviation of your Normal random variable?",
          value = "0",
          width = '80%', placeholder = NULL)

textInput("x1",
          "What cutoff value do you want to find the area below?",
          value = "0",
          width = '80%', placeholder = NULL)

bs_1 = reactive("pnorm(%ni, mean = %ni, sd = %ni)")

```

```{r, context="server"}

output$code_1 <- renderText(
        makePrintText(
          base_string = bs_1(),
          old_input = c(),
          new_input = c(input$x1, input$mu1, input$sigma1)
        )
  )

output$output_1 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
          base_string = bs_1(),
          old_input = c(),
          new_input = c(input$x1, input$mu1, input$sigma1)
    )
   ))
})

output$norm_var_1 <- renderText(
        makePrintText(
          base_string = "X ~ Normal(%ni, %ni) \n P(X ≤ %ni) = ?",
          old_input = c(),
          new_input = c(input$mu1, input$sigma1, input$x1)
        )
  )


output$distn_1 <- renderPlot(
  illustrate_norm(input$mu1, input$sigma1, x = input$x1)
)

```

Random variable and probability statement:
```{r}
htmlOutput("norm_var_1")
```

Code:
```{r}
htmlOutput("code_1")
```

Result from code:
```{r}
verbatimTextOutput("output_1")
```


Illustration:
```{r}
plotOutput("distn_1")
```

### Your turn

In the box below, you may run any line of **R** code.  Use the box, and/or the exercise above, to answer the following questions.


```{r ex_1.1, exercise = TRUE}

```

Suppose heights of people are Normally distributed, with a mean of 67 inches and a standard deviation of 3 inches.  Use the code box below to calculate your answers to the following questions.  (DRAW PICTURES!)

*Round your answers to 2 decimals*

```{r q_1}
textInput("a_1.2",
           "What is the probability that a randomly sampled person is less than 5 feet (60 inches) tall?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.2")

textInput("a_1.1",
           "What is the probability that a randomly sampled person is at least 71 inches tall?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.1")

textInput("a_1.3",
           "What is the area below a z-score of -1.38?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_1.3")


```

```{r, context = "server"}

output$check_1.1 <- renderText(
  if(input$a_1.1 == ""){
     ""
  }else if(round(as.numeric(input$a_1.1),2) == round(1-pnorm(71, mean = 67, sd = 3), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.2 <- renderText(
  if(input$a_1.2 == ""){
     ""
  }else if(round(as.numeric(input$a_1.2), 2) == round(pnorm(60, mean = 67, sd = 3), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_1.3 <- renderText(
  if(input$a_1.3 == ""){
     ""
  }else if(round(as.numeric(input$a_1.3), 2) == round(pnorm(-1.38, mean = 0, sd = 1), 2)){
      correct
  }else{
    incorrect
  }
)
```



## Finding values from probabilities

As we saw with the Binomial exercise, sometimes we want to ask probability questions "in reverse".  For example, we might ask "What SAT score puts someone in the 95th percentile of all scores?". (That is, what score is better than 95% of all scores.)  We could answer this via guess-and-check with `pnorm()`, but fortunately, there is a function for it in **R** called `qnorm()`.


```{r qnorm}
textInput("mu2",
          "What is the mean of your Normal random variable?",
          value = "0",
          width = '80%', placeholder = NULL)


textInput("sigma2",
          "What is the standard deviation of your Normal random variable?",
          value = "0",
          width = '80%', placeholder = NULL)

textInput("q2",
          "We want to find the value that has an area of ____ below it.",
          value = "0",
          width = '80%', placeholder = NULL)

bs_3 = reactive("qnorm(%ni, mean = %ni, sd = %ni)")

```

```{r, context="server"}

output$code_3 <- renderText(
        makePrintText(
          base_string = bs_3(),
          old_input = c(),
          new_input = c(input$q2, input$mu2, input$sigma2)
        )
  )

output$output_3 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_3(), 
      old_input = c(),
      new_input = c(input$q2, input$mu2, input$sigma2)
    )
   ))
})

output$norm_var_3 <- renderText(
        makePrintText(
          base_string = "X ~ Normal(%ni, %ni) \n P(X ≤ ?) = %ni",
          old_input = c(),
          new_input = c(input$mu2, input$sigma2, input$q2)
        )
  )


output$distn_3 <- renderPlot(
  illustrate_norm(input$mu2, input$sigma2, q = input$q2)
)

```

Random variable and probability statement:
```{r}
htmlOutput("norm_var_3")
```

Code:
```{r}
htmlOutput("code_3")
```

Result:
```{r}
verbatimTextOutput("output_3")
```


Illustration:
```{r}
plotOutput("distn_3")
```

### Your turn

Suppose SAT scores are Normally distributed with a meand of 1600 and a standard deviation of 200.

```{r ex_3, exercise = TRUE}

```

```{r q_3}
textInput("a_3.1",
           "What SAT score would put someone in the 95th percentile?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.1")

textInput("a_3.2",
           "What SAT score is worse than 80% of all scores?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.2")

textInput("a_3.3",
           "What is the third quartile of SAT scores?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.3")

textInput("a_3.4",
           "What z-score has 95% below it?",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_3.4")

```

```{r, context = "server"}

output$check_3.1 <- renderText(
  if(input$a_3.1 == ""){
     ""
  }else if(round(as.numeric(input$a_3.1),2) == round(qnorm(.95, 1600, 200), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_3.2 <- renderText(
  if(input$a_3.2 == ""){
     ""
  }else if(round(as.numeric(input$a_3.2),2) == round(qnorm(.20, 1600, 200), 2)){
      correct
  }else if(round(as.numeric(input$a_3.2),2) == round(qnorm(.20, 1600, 200), 2)){
    "<font color='red'>Not quite... draw a picture!</font><br><br>"
  }else{
    incorrect
  }
)

output$check_3.3 <- renderText(
  if(input$a_3.3 == ""){
     ""
  }else if(round(as.numeric(input$a_3.3),2) == round(qnorm(.75, 1600, 200), 2)){
      correct
  }else{
    incorrect
  }
)

output$check_3.4 <- renderText(
  if(input$a_3.4 == ""){
     ""
  }else if(round(as.numeric(input$a_3.4),2) == round(qnorm(.95, 0, 1), 2)){
      correct
  }else{
    incorrect
  }
)
```



## Challenge:  Practice predicting code

A good habit to get into when coding is to "translate" your code to English in your head, and try to predict what the output will be.  This is how to figure out if you are truly asking the computer to calculate what you hoped.  (Remember - computers will almost always output *something*!  It is your responsibility as the scientist to make sure you have asked the right question.)

For the following, try to guess what the output would be.  Do **not** run the code; instead, try to read it as a probability question and think through the answer.

Hint: **DRAW A PICTURE** for each question!


*Give answers to 2 decimals*


```{r q_4}
textInput("a_4.1",
           "pnorm(0, mean = 0, sd = 1)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_4.1")

textInput("a_4.2",
           "pnorm(-1, mean = 0, sd = 1)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_4.2")

textInput("a_4.3",
           "pnorm(7, mean = 5, sd = 2)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_4.3")


textInput("a_4.4",
           "qnorm(0.975, mean = 5, sd = 2)",
           value = "",
           width = '80%', placeholder = NULL)

htmlOutput("check_4.4")

```

```{r, context = "server"}

output$check_4.1 <- renderText(
  if(input$a_4.1 == ""){
     ""
  }else if(as.numeric(input$a_4.1) == 0.5){
      correct
  }else{
    incorrect
  }
)

output$check_4.2 <- renderText(
   if(input$a_4.2 == ""){
     ""
  }else if(as.numeric(input$a_4.2) == 0.16){
      correct
  }else{
    incorrect
  }
)

output$check_4.3 <- renderText(
  if(input$a_4.3 == ""){
     ""
  }else if(as.numeric(input$a_4.3) == 0.84){
      correct
  }else{
    incorrect
  }
)

output$check_4.4 <- renderText(
  if(input$a_4.4 == ""){
     ""
  }else if(as.numeric(input$a_4.4) == 9){
      correct
  }else{
    incorrect
  }
)

```
